### 2020-08-10 analysis/analysis_stable.Rmd
Adding the analysis notebook.

---
title: "analysis_stable"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
date: "`r substr(Sys.time(), 1, 10)`"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  fig.align = "center",
  fig.retina = 2,
  fig.width = 10,
  cache.lazy = FALSE
  )

library(tidyverse)
library(stringr)
library(viridis)
library(vctrs)

library(DT)

library(fst)

library(bench)

source("insights.R", local = knitr::knit_global())
```

All insights about eval calls are to put here, after the computations are stabilized (so use another notebook to explore).

# Setup

```{r eval_calls_file}
eval_calls_file <- "/var/lib/R/project-evalR/run/package-evals-traced-all.1/calls.fst"
```

```{r load_file}
eval_calls <- read_fst(eval_calls_file)
eval_calls <- as_tibble(eval_calls) 
```

```{r resolve_types}
eval_calls <- eval_calls %>% resolve_sexp_name(expr_expression_type) %>% 
                             resolve_sexp_name(expr_resolved_type) %>%
                             resolve_sexp_name(enclos_type) %>%
                             resolve_sexp_name(envir_type)
```

# Usage metrics

## R code size

## Number of `eval`

For the number of call sites, we either use the `srcref` when there is one. When there is not (such as in *base*), we use a combination of file and `eval_expr_expression`.

```{r}
nb_call_sites <- eval_calls %>% select(eval_call_srcref) %>% n_distinct(na.rm = TRUE)  + eval_calls %>% filter(is.na(eval_call_srcref)) %>% select(file, eval_call_expression)%>% n_distinct()
```

There are `r nb_call_sites` different `eval` call sites.

## Number of calls to `eval`

```{r}
nb_eval_calls <- eval_calls %>% nrow()
```

There are `r nb_eval_calls` eval calls.

## Size of expressions as arguments to `eval`

## Amount of computations performed in `eval`

## Aliases of `eval`

# A taxonomy of `eval`

## The operation mix

## Scope (environments)

## Patterns

## Provenance

## Consistence

# Other faces of `eval`

# Case studies