### 2020-08-10 main.tex
Added some comments to the paper

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{On the Design, Use and Removal of Eval in R}

\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
}

\maketitle

\begin{abstract}
This document is a model and instructions for \LaTeX.
This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
or Math in Paper Title or Abstract.
\end{abstract}

\begin{IEEEkeywords}
component, formatting, style, styling, insert
\end{IEEEkeywords}

\section{Introduction}
- Why eval is used in R
- What the community says about using eval
- What existing tools do about eval
- Do people use eval for obvious things like looking up variable, if yes, it is
because they are unaware of alternatives?
- Why automating the transformation is a hard problem.
- What is the goal of the work? To replace eval with code for static analyssi or
for consumption by the users of the language?
- We only permit transformations that preserve code behavior?
- Do we accept overly complex code as the output?

Why do we do this dynamically? Because it raises a chicken-and-egg problem,
doing it statically would require us to reason about eval based code which may
not be possible in general.

What are the challenges? Why is it not trivial?
- Do we handle aliasing of eval?
- Simple cases can be subtle? (laziness for ex.

Why can't we replace eval(code) by code?

Intended user of the technique? Which version of R is targeted.

Contributions:
- Describe the use of eval in R
- Develop a tool to replace evals guided by dynamic traces

- cannot replace with (cite study of with construct).
\textbf{*}

\section{Background on R}
\textbf{Aviral}

\section{The Design of Eval}
\textbf{Aviral}

\section{Infrastructure}
\textbf{Aviral + Filip}

\section{Corpus of R Programs}
\textbf{Filip}

\section{Threats to Validity}
\textbf{Filip}

\section{Analyzing Eval Usage Patterns}
This section presents the results of our empirical study of eval in the R language.
\subsection{How many evals}
\textbf{Pierre}

\subsection{How are they used}
\textbf{Pierre}
- classify evals based on what they can be replaced with

\subsection{How to replace evals}
\textbf{Aviral}

\subsection{How many evals can be replaced}
\textbf{Aviral}

\section{Evaluation}

\section{Related Work}
\eval has been widely discussed in the research literature for the challenges it
imposes for code comprehension and automated code analyses.
Perhaps, \citeit{EvalMenDo} is the most popular exposition of \eval and its
challenges for JavaScript.

\textbf{Aviral}
- javascript
- ruby
- php

\section{Conclusion}
\textbf{*}

\section*{Acknowledgment}

\section*{References}

\end{document}