### 2020-08-21 inst/analysis/trace-run-analysis.Rmd
Add trace run analysis

---
title: "Trace run analysis"
output: 
  html_document: 
    theme: united
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: console
params:
  data_dir: /var/lib/R/project-evalR/run/package-evals-traced-corpus.3
  output_dir: /var/lib/R/project-evalR/revalstudy/data
  paper_dir: /var/lib/R/project-evalR/revalstudy/inst/paper
  base_url: http://prl3:8003/file_show?path=
---

```{r setup, include=FALSE}
for (x in c(
  "dplyr", 
  "DT", 
  "fs", 
  "purrr",
  "readr",
  "runr",
  "stringr"
  )) {
  suppressPackageStartupMessages(library(x, character.only=TRUE))
}

knitr::opts_chunk$set(echo = TRUE)

# TODO: move into the package
source("inc/paths.R")
source("inc/setup.R")
source("inc/latextags.R")

create_tags(path(TAGS_DIR, "trace-run.tex"), prefix="TC", default=TRUE)

PARALLEL_LOG_FILE <- path(params$data_dir, "parallel.log")
```

## Description

The aim of this notebook is to analyze the run of the tracer.

### Output

- ``r PARALLEL_LOG_FILE`` - GNU parallel log file

## Load data

```{r}
parallel_log <- 
  read_parallel_log(PARALLEL_LOG_FILE) %>%
  mutate(command=str_replace(command, ".*/(.*)$", "\\1")) %>%
  select(package=command, everything()) %>%
  select(-seq, -host, -send, -receive)
```

## Overview

```{r}
failed_packages <- filter(parallel_log, exitval != 0)
errored_packages <- filter(parallel_log, exitval == 1)
timeouted_packages <- filter(parallel_log, exitval == -1)

overview_table(
  r("Run packages", parallel_log),
  r("Succeeded packages", filter(parallel_log, exitval == 0)),
  r("Failed packages",    failed_packages),
  r("Failed packages ratio", ratio(failed_packages, parallel_log)),
  r("Timeouted packages", timeouted_packages),
  r("Errored packages", errored_packages),
  r("Other errored packages",   filter(parallel_log, !(exitval %in% c(-1, 0, 1)))),
  r("Duration", max(parallel_log$starttime + parallel_log$jobruntime) - min(parallel_log$starttime))
)
```

## Failed packages

### Errored packages

Running tracer on these packages resulted in exit code `1`.

```{r}
errored_packages %>%
  mutate(
    stdout=paste("<a href=", URLencode(file.path(params$base_url, params$data_dir, package, "stdout")), ">stdout</a>"),
    stderr=paste("<a href=", URLencode(file.path(params$base_url, params$data_dir, package, "stderr")), ">stderr</a>")
  ) %>%
  my_datatable(escape=FALSE)
```

### Timeouted packages

The GNU parallel killed the tracing process.

```{r}
timeouted_packages %>%
  my_datatable()
```