### 2020-08-23 inst/analysis/kaggle.Rmd

---
title: "Kaggle"
output: html_document
params:
  base_dir: /home/rstudio/evalR
  hostname: prl2
  port: 8788
  competitions:
    - titanic
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
for (x in c(
  "dplyr", 
  "DT", 
  "fs",
  "fst",
  "purrr",
  "readr",
  "rjson",
  "runr",
  "stringr",
  "tibble"
  )) {
  suppressPackageStartupMessages(library(x, character.only=TRUE))
}

knitr::opts_chunk$set(echo = TRUE)

# TODO: move into the package
source("inc/paths.R")
source("inc/setup.R")
source("inc/latextags.R")

create_tags(path(TAGS_DIR, "kaggle.tex"), prefix="Kaggle", default=TRUE)

KAGGLE_KERNELS_DIR  <- path(params$base_dir, "run", "kaggle-kernels", "notebooks", "r", "kernels")
KAGGLE_DATASETS_DIR <- path(params$base_dir, "run", "kaggle-datasets")
KAGGLE_RUN_DIR      <- path(params$base_dir, "run", "kaggle-run")

KERNELS_FILE      <- path(KAGGLE_RUN_DIR, "kernels.csv")
SCRIPTS_FILE      <- path(KAGGLE_RUN_DIR, "scripts.txt")
PARALLEL_LOG_FILE <- path(KAGGLE_RUN_DIR, "parallel.csv")
```

```{r check}
stopifnot(dir_exists(KAGGLE_KERNELS_DIR))
stopifnot(dir_exists(path(KAGGLE_DATASETS_DIR, params$competitions)))
```

```{r aux functions}
wrap <- function(file, body) {
  str_glue(
    "evil::write_eval_traces(",
    "  evil::trace_eval(",
    "    quote = TRUE,",
    "    code = {{",
    "      {gsub('\n', '\n      ', body)}",
    "    }}",
    "  ),",
    "  datadir=getwd()",
    ")",
    .sep = "\n"
  )
}

process <- function(x) {
  run_path <- path(KAGGLE_RUN_DIR, x$competition, basename(x$path))
  code_file <- path(x$path, "script", x$code_file)
  run_code_file <- path(run_path, "run.R")

  dir_create(run_path)

  if (str_ends(code_file, "\\.[rR]$")) {
    file_copy(code_file, run_code_file, overwrite=TRUE)
  } else if (str_ends(code_file, "\\.Rmd")) {
    knitr::purl(code_file, run_code_file, quiet=TRUE)
  } else if (str_ends(code_file, "\\.irnb") || str_ends(code_file, "\\.ipynb")) {
    tmp <- tempfile(fileext = "Rmd")
    rmarkdown:::convert_ipynb(code_file, tmp)
    knitr::purl(tmp, run_code_file, quiet=TRUE)
  }

  if (file_exists(run_code_file)) {
    tmp <- read_file(run_code_file)
    tmp <- wrap(run_code_file, tmp)
    write_file(tmp, run_code_file)
  }

  run_code_file
}
```

```{r load kaggle metadata}
files <- list.files(KAGGLE_KERNELS_DIR, "kernel-metadata\\.json$", full.names=TRUE, recursive=TRUE)

metadata_json <- map(files, ~fromJSON(file = .))
metadata <- map_dfr(
  metadata_json, 
  ~tibble(id=str_replace(.$id, "/", "-"), language=.$language, kernel_type=.$kernel_type, competition=.$competition_sources, code_file=.$code_file)
)

kernels <- 
  metadata %>%
  mutate(
    path=path(KAGGLE_KERNELS_DIR, id)
  ) %>% 
  filter(dir_exists(path))
```

```{r filter known competitions}
known_competitions <- semi_join(kernels, tibble(competition=params$competitions), by="competition")
```

```{r delete run dir}
if (dir_exists(KAGGLE_RUN_DIR)) dir_delete(KAGGLE_RUN_DIR)
```

```{r extract kernels}
extraction <- 
  known_competitions %>%
  rowwise() %>%
  do({
    tryCatch({
      tibble(id=.$id, run_file=process(.))
    }, error=function(e) {
      tibble(id=.$id, error=e$message)
    })
  })
```

## Prepare for run

```{r make a kernels_supported set}
kernels_supported <- 
  known_competitions %>%
  left_join(extraction, by="id") %>%
  filter(file_exists(run_file))
```

```{r compute sloc}
sloc <- map_dfr(path(KAGGLE_RUN_DIR, params$competitions), function(x) {
  out <- system2("cloc", c("--by-file-by-lang", "-q", "--csv", shQuote(x)), stdout = TRUE)  
  out <- out[2:(length(out)-4)]
  out[1] <- "language,filename,blank,comment,code"
  df <- read_csv(out)
}) %>%
  mutate(
    kernel=basename(dirname(filename))
  ) %>%
  select(kernel, code)

stopifnot(any(!duplicated(sloc$kernel)))
```

```{r add sloc to the kernels_supported set}
kernels_supported <- left_join(kernels_supported, sloc, by=c("id"="kernel"))
```

```{r kernels with no R code or duplication}
kernels_duplicated <- filter(kernels_supported, is.na(code) | code == 0)
kernels_runnable <- filter(kernels_supported, code > 0)
```

```{r display duplicated or empty kernels}
kernels_duplicated %>% 
  mutate(
    code_file=show_url(file.path(KAGGLE_KERNELS_DIR, id, "script", code_file)),
    path=show_url(path),
    run_file=show_url(run_file)
  ) %>%
  select(id, competition, path, run_file, language, code_file) %>%
  my_datatable(escape=FALSE)
```

```{r}
write_csv(kernels_runnable, KERNELS_FILE)
write_lines(dirname(kernels_runnable$run_file), SCRIPTS_FILE)
```

## Run

```{r run}
Sys.unsetenv("R_HOME")
run <- system2(
  "parallel",
  c(
    "-a", SCRIPTS_FILE,
    "--bar",
    "--jobs", parallel::detectCores(),
    "--results", PARALLEL_LOG_FILE,
    "--tagstring", "{/}",
    "--timeout", "1h",
    "--workdir", "{1}/",
    file.path(R_DIR, "bin", "R"), "CMD", "BATCH", "{1}/run.R", "{1}/run.R.out"
  )
)
```

```{r load results}
calls <- kernels_runnable %>%
  do({
    bd <- basename(.$run_file)
    tryCatch({
      df <- as_tibble(read_fst(path(bd, "calls.fst")))
      add_column(df, id=.$id, .before=1)
    }, error=function(e) {
      tibble(id=.$id, error=e$message)
    })
  })
```

```{r}
parallel_log <- 
  read_csv(PARALLEL_LOG_FILE)

parallel_log %>%
  mutate(
    output=show_url(path(V1, "run.R.out"))
  ) %>%
  my_datatable(escape=FALSE)
```

## Errors