### 2020-08-23 inst/analysis/kaggle.Rmd

---
title: "Kaggle"
output: html_document
params:
  base_dir: /var/lib/R/project-evalR
  hostname: prl3
  port: 8788
  competitions:
    - titanic
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
for (x in c(
  "callr",
  "dplyr", 
  "DT", 
  "fs",
  "fst",
  "knitr",
  "purrr",
  "readr",
  "rjson",
  "rmarkdown",
  "runr",
  "stringr",
  "tibble"
  )) {
  suppressPackageStartupMessages(library(x, character.only=TRUE))
}

knitr::opts_chunk$set(echo = TRUE)

# TODO: move into the package
source("inc/paths.R")
source("inc/setup.R")
source("inc/latextags.R")
```

```{r consts}
KAGGLE_KERNELS_DIR  <- path(RUN_DIR, "kaggle-kernels", "notebooks", "r", "kernels")
KAGGLE_DATASETS_DIR <- path(RUN_DIR, "kaggle-datasets")
KAGGLE_RUN_DIR      <- path(RUN_DIR, "kaggle-run")

KERNELS_FILE      <- path(KAGGLE_RUN_DIR, "kernels.csv")
SCRIPTS_FILE      <- path(KAGGLE_RUN_DIR, "scripts.txt")
PARALLEL_LOG_FILE <- path(KAGGLE_RUN_DIR, "parallel.log")
```

```{r check}
stopifnot(dir_exists(KAGGLE_KERNELS_DIR))
stopifnot(dir_exists(path(KAGGLE_DATASETS_DIR, params$competitions)))
```

```{r aux functions}
wrap <- function(file, body) {
  str_glue(
    "evil::write_eval_traces(",
    "  evil::trace_eval(",
    "    quote = TRUE,",
    "    code = {{",
    "      {gsub('\n', '\n      ', body)}",
    "    }}",
    "  ),",
    "  datadir=getwd()",
    ")",
    .sep = "\n"
  )
}

extract_kernel_code <- function(path, code_file, competition) {
  run_path <- path(KAGGLE_RUN_DIR, competition, basename(path))
  code_file <- path(path, "script", code_file)
  run_code_file <- path(run_path, "run.R")

  if (!dir_exists(run_path)) dir_create(run_path)

  if (str_ends(code_file, "\\.[rR]$")) {
    file_copy(code_file, run_code_file, overwrite=TRUE)
  } else if (str_ends(code_file, "\\.Rmd")) {
    knitr::purl(code_file, run_code_file, quiet=TRUE)
  } else if (str_ends(code_file, "\\.irnb") || str_ends(code_file, "\\.ipynb")) {
    tmp <- tempfile(fileext = "Rmd")
    rmarkdown:::convert_ipynb(code_file, tmp)
    knitr::purl(tmp, run_code_file, quiet=TRUE)
  }

  if (file_exists(run_code_file)) {
    tmp <- read_file(run_code_file)
    tmp <- wrap(run_code_file, tmp)
    write_file(tmp, run_code_file)
  }

  run_code_file
}
```

## Description

This notebook prepares the kaggle kernels to be run with the eval tracing framework.

### Input

- `r KAGGLE_DATASETS_DIR` - directory with kaggle datasets
- `r KAGGLE_KERNELS_DIR` - directory with kaggle kernels

### Output

- `r KAGGLE_RUN_DIR` - directory with prepared kernels to be run
- command to run the datasets

### Load kernels

```{r load kaggle metadata}
files <- list.files(KAGGLE_KERNELS_DIR, "kernel-metadata\\.json$", full.names=TRUE, recursive=TRUE)

metadata_json <- map(files, ~fromJSON(file = .))
metadata <- map_dfr(
  metadata_json, 
  ~tibble(id=str_replace(.$id, "/", "-"), language=.$language, kernel_type=.$kernel_type, competition=.$competition_sources, code_file=.$code_file)
)

kernels <- 
  metadata %>%
  mutate(
    path=path(KAGGLE_KERNELS_DIR, id)
  ) %>% 
  filter(dir_exists(path))
```

```{r filter known competitions}
known_competitions <- semi_join(kernels, tibble(competition=params$competitions), by="competition")
```

```{r delete run dir}
if (dir_exists(KAGGLE_RUN_DIR)) {
  warning("*** ", KAGGLE_RUN_DIR, " exists")
}
```

```{r extract kernels}
extraction_lst <- pbapply::pbapply(known_competitions, 1, cl=16, function(x) {
    tryCatch({
      tibble(id=x["id"], run_file=extract_kernel_code(x["path"], x["code_file"], x["competition"]))
    }, error=function(e) {
      tibble(id=x["id"], error=e$message)
    })
})
extraction <- bind_rows(extraction_lst)
```

## Prepare for run

```{r make a kernels_supported set}
kernels_supported <- 
  known_competitions %>%
  left_join(extraction, by="id") %>%
  filter(file_exists(run_file))
```

```{r compute sloc}
sloc <- map_dfr(path(KAGGLE_RUN_DIR, params$competitions), function(x) {
  out <- system2("cloc", c("--include-lang=R", "--by-file-by-lang", "-q", "--csv", shQuote(x)), stdout = TRUE)  
  out_r <- out[startsWith(out, "R,")]
  out_csv <- c("language,filename,blank,comment,code", out_r)
  df <- read_csv(out_csv)
}) %>%
  mutate(
    kernel=basename(dirname(filename))
  ) %>%
  select(kernel, code)

stopifnot(any(!duplicated(sloc$kernel)))
```

```{r add sloc to the kernels_supported set}
kernels_supported <- 
  kernels_supported %>%
  left_join(sloc, by=c("id"="kernel")) %>%
  mutate(runnable=ifelse(is.na(code), FALSE, code > 0))
```

```{r display duplicated or empty kernels}
filter(kernels_supported, !runnable) %>% 
  mutate(
    code_file=show_url(file.path(KAGGLE_KERNELS_DIR, id, "script", code_file)),
    path=show_url(path),
    run_file=show_url(run_file)
  ) %>%
  select(id, competition, path, code, run_file, language, code_file) %>%
  my_datatable(escape=FALSE)
```

```{r}
kernels_runnable <- filter(kernels_supported, runnable)
write_csv(kernels_supported, KERNELS_FILE)
write_lines(dirname(kernels_runnable$run_file), SCRIPTS_FILE)
```

```{r}
count(kernels_supported, runnable)
```

## Run

Run the following command from the `r params$base_dir` directory:

```{r}
cat("cd", shQuote(params$base_dir))

cat("parallel",
    "-a", shQuote(SCRIPTS_FILE),
    "--bar",
    "--jobs", parallel::detectCores(),
    "--results", shQuote(PARALLEL_LOG_FILE),
    "--tagstring", shQuote("{/}"),
    "--timeout", "1h",
    "--workdir", shQuote("{1}/"),
    path(R_DIR, "bin", "R"), "CMD", "BATCH", shQuote("{1}/run.R"), shQuote("{1}/run.R.out"), 
    "\n"
)

cat("runr/inst/merge-fst.R", "run/kaggle-run", "calls.fst", "ERROR.fst", "\n")
```