### 2020-08-23 inst/analysis/corpus-analysis.Rmd
First paragraph in 5

---
title: "Corpus Analysis"
output: html_document
params:
  base_dir: ../../../
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
for (x in c(
  "dplyr", 
  "DT",
  "fs",
  "fst",
  "ggplot2",
  "ggthemes",
  "lubridate",
  "readr",
  "purrr",
  "stringr",
  "tidyr"
  )) {
  suppressPackageStartupMessages(library(x, character.only=TRUE))
}

knitr::opts_chunk$set(
  echo = TRUE,
  fig.align = "center",
  fig.retina = 2,
  fig.width = 10
)

# TODO: move into the package
source("inc/paths.R")
source("inc/setup.R")
source("inc/latextags.R")

options(repos="https://cran.r-project.org")
theme_set(theme_minimal())
create_tags(path(TAGS_DIR, "corpus.tex"), prefix="Corpus", default=TRUE)
```

## Load data

```{r}
corpus <- as_tibble(read_fst(CORPUS_DETAILS_FILE))
```

## Corpus Overview

### Table

```{r summary}
overview_table(
  r("packages",             corpus),
  r("runnable code",        sum(corpus$runnable_code, na.rm=T)),
  r("examples code",        sum(corpus$runnable_code_examples, na.rm=T)),
  r("tests code",           sum(corpus$runnable_code_tests, na.rm=T)),
  r("vignettes code",       sum(corpus$runnable_code_vignettes , na.rm=T)),
  r("mean expr coverage",   percent(mean(corpus$coverage_expr))),
  r("median expr coverage", percent(median(corpus$coverage_expr))),
  r("with R code",          filter(corpus, package_r_code > 0)),
  r("R code",               sum(corpus$package_r_code, na.rm=T)),
  r("with native code",     filter(corpus, package_native_code > 0)),
  r("native code",          sum(corpus$package_native_code, na.rm=T)),
  r("functions",            sum(corpus$funs_public+corpus$funs_private, na.rm=T)),
  r("public functions",     sum(corpus$funs_public, na.rm=T)),
  r("private functions",    sum(corpus$funs_private, na.rm=T)),
  r("ranked in cranlogs",   filter(corpus, !is.na(rank))),
  r("mean revdes",          mean(corpus$revdeps)),
  r("median revdes",        median(corpus$revdeps)),
  r("runtime",              as.duration(sum(corpus$runnable_time))),
  r("mean runtime",         as.duration(mean(corpus$runnable_time))),
  r("median runtime",       as.duration(median(corpus$runnable_time))),
  r("function with eval",   sum(corpus$funs_with_eval, na.rm=T))
)
```

### CRAN

```{r}
overview_table(
  r("all cran", dim(available.packages())[1])
)
```

### Plots

```{r plot corpus}
coverage_mean <- mean(corpus$coverage_expr)
coverage_median <- median(corpus$coverage_expr)
revdeps_mean <- mean(corpus$revdeps)
revdeps_median <- median(corpus$revdeps)

corpus %>%
  mutate(
    outlier=package_code > PACKAGE_SIZE_OUTLIER,
    label=ifelse(rank <= 25 | outlier, package, NA)  #if_else(outlier, str_glue("{package}"), as.character(NA))
  ) %>%
  ggplot(
    aes(
      x=coverage_expr, 
      y=revdeps,
      label=label,
      size=package_code
    )
  ) +
  geom_point(alpha=.6) +
  geom_text(size=2, check_overlap = T, vjust=1.5, na.rm = TRUE) +

  geom_hline(aes(yintercept=revdeps_mean), linetype="dashed", color="red", size=0.2) +
  geom_vline(aes(xintercept=coverage_mean), linetype="dashed", color="red", size=0.2) +

  scale_x_continuous(labels = scales::percent) + 
  scale_y_log10(labels = scales::comma) +
  scale_size(range = c(.1, 3), labels = scales::comma) +
  theme(
    legend.position=c(0.2, 0.85),
    legend.box="horizontal",
    legend.box.background = element_rect(fill="white", size=0.1)
  ) +
  labs(
    x="Code coverage", 
    y="Number of reverse dependencies (log)", 
    size="Code size (R + native)",
    linetype=""
  ) +
  guides(
    size=guide_legend(order=1),
    color=guide_legend(order=2),
    linetype=F
  )

ggsave(CORPUS_PLOT)
```

### List of corpus packages

```{r list of packages, eval=FALSE}
datatable(
  corpus, 
  filter="top", 
  options=list(
    pageLength=50,
    autoWidth=TRUE
  )
)
```

## Kaggle

```{r, eval=FALSE}
parallel_log <- 
  read_csv(PARALLEL_LOG_FILE)

parallel_log %>%
  mutate(
    output=show_url(path(V1, "run.R.out"))
  ) %>%
  my_datatable(escape=FALSE)
```