### 2020-08-25 inst/analysis/corpus-analysis.Rmd
V.A

---
title: "Corpus Analysis"
output: html_document
params:
  base_dir: ../../../
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
local(for (x in c(
  "dplyr", 
  "DT",
  "fs",
  "fst",
  "ggplot2",
  "ggthemes",
  "lubridate",
  "readr",
  "purrr",
  "stringr",
  "tidyr",
  "viridis"
  )) {
  suppressPackageStartupMessages(library(x, character.only=TRUE))
})

knitr::opts_chunk$set(
  echo = TRUE,
  fig.align = "center",
  fig.retina = 2,
  fig.width = 10
)

# TODO: move into the package
source("inc/paths.R")
source("inc/setup.R")
source("inc/latextags.R")

options(repos="https://cran.r-project.org")
theme_set(theme_minimal())
create_tags(path(TAGS_DIR, "corpus.tex"), prefix="Corpus", default=TRUE)
```

## Load data

```{r load corpus}
corpus <- 
  read_fst(CORPUS_DETAILS_FILE) %>%
  as_tibble() %>%
  mutate_at(
    vars(
      starts_with("runnable_code"),
      starts_with("runnable_files"),
      starts_with("package_"),
      starts_with("funs_"),
      "evals"
    ), 
    replace_na, 
    0
  )
```

```{r load kaggle}
kaggle_kernels <- as_tibble(read_fst(KAGGLE_KERNELS_FILE))
kaggle_log <- as_tibble(read_fst(KAGGLE_LOG_FILE)) %>%
  transmute(
    id=str_replace(Command, ".*/kaggle-run/[^/]+/([^/]+)/.*", "\\1"),
    exitval=Exitval,
    runtime=as.duration(JobRuntime)
  )

kaggle <- left_join(
  kaggle_kernels,
  kaggle_log,
  by="id"
)
```

```{r load runnable code metadata}
runnable_code_metadata <- read_fst(PACKAGE_RUNNABLE_CODE_FILE) %>%
  as_tibble() %>%
  mutate(package=basename(package)) %>%
  select(package, file=filename, code)

# TODO semi_join with traced-programs.fst
```

```{r load run metadata}
runs <- read_fst(PACKAGE_RUNS_FILE) %>%
  as_tibble() %>%
  mutate(
    package=basename(package),
    file=as.character(path("./", basename(dirname(file)), basename(file)))
  ) %>%
  semi_join(corpus, by="package")
```

## CRAN packages

```{r compute corpus programs}
corpus_programs <-
  left_join(
    filter(runs, exitval==0) %>% select(package, file, time), 
    runnable_code_metadata %>% select(package, file, code),
    by=c("package", "file")
  ) %>%
  transmute(
    package,
    file,
    type=basename(dirname(file)),
    code, 
    runtime=time
  )
```

There will be some programs that were not strictly part of examples, tests or vignettes, or they do not have
counts for code size or for execution time.

```{r compute program inconsistencies}
corpus_programs_inconsistent <-
  corpus_programs %>%
  filter(
    !(type %in% c("examples", "tests", "vignettes")) |
    is.na(code) | 
    is.na(runtime) |
    code == 0
  )

corpus_programs_inconsistent %>% 
    my_datatable()
```

Since, there are just a few of them, we ignore them.

```{r filter out inconsistent programs}
corpus_programs <- anti_join(corpus_programs, corpus_programs_inconsistent, by=c("package", "file"))
```

```{r summary}
runnable_code <- 
  corpus_programs %>%
  mutate(files=1) %>%
  pivot_wider(names_from=type, values_from=c(files, code)) %>%
  select(-file) %>%
  group_by(package) %>%
  summarise_all(sum, na.rm=TRUE) %>%
  group_by(package) %>%
  mutate(
    files=files_examples + files_tests + files_vignettes,
    code=code_examples + code_tests + code_vignettes
  )

rs <- function(name, value) {
  stopifnot(is.numeric(value))
  bind_rows(
    r(name, sum(value)),
    r(paste(name, "mean"), mean(value)),
    r(paste(name, "median"), median(value)),
    r(paste(name, "min"), min(value)),
    r(paste(name, "max"), max(value))
  )
}

corpus_program_examples <- filter(corpus_programs, type=="examples")
corpus_programs_tests <- filter(corpus_programs, type=="tests")
corpus_programs_vignettes <- filter(corpus_programs, type=="vignettes")

overview_table(
  r("packages",    corpus),
  rs("revdeps", corpus$revdeps),
  r("in cranlogs", filter(corpus, !is.na(rank))),

  rs("package programs",         runnable_code$files),
  rs("package programs code",    runnable_code$code),
  rs("package programs runtime", as.duration(runnable_code$runtime)),

  rs("examples programs",  runnable_code$files_examples),
  rs("tests programs",     runnable_code$files_tests),
  rs("vignettes programs", runnable_code$files_vignettes),

  rs("examples code",  corpus_program_examples$code),
  rs("tests code",     corpus_programs_tests$code),
  rs("vignettes code", corpus_programs_vignettes$code),

  rs("examples runtime",  corpus_program_examples$runtime),
  rs("tests runtime",     corpus_programs_tests$runtime),
  rs("vignettes runtime", corpus_programs_vignettes$runtime),

  r("mean expr coverage",   percent(mean(corpus$coverage_expr))),
  r("median expr coverage", percent(median(corpus$coverage_expr))),

  r("with R code",      filter(corpus, package_r_code > 0)),
  r("with native code", filter(corpus, package_native_code > 0)),

  rs("code",        corpus$package_code),
  rs("R code",      corpus$package_r_code),
  rs("native code", corpus$package_native_code),

  rs("functions",          corpus$funs_public + corpus$funs_private),
  rs("function with eval", corpus$funs_with_eval),
  rs("public functions",   corpus$funs_public),
  rs("private functions",  corpus$funs_private),

  rs("revdes", corpus$revdeps)
)
```

```{r generals}
overview_table(
  r("all cran", dim(available.packages())[1]),
  r("core packages", length(CORE_PACKAGES)),
  r("most downloaded treshold", MOST_DOWNLOADED),
  r("large package treshold", PACKAGE_SIZE_OUTLIER)
)
```

### Plot

```{r plot corpus}
coverage_mean   <- mean(corpus$coverage_expr)
coverage_median <- median(corpus$coverage_expr)
revdeps_mean    <- mean(corpus$revdeps)
revdeps_median  <- median(corpus$revdeps)
funs_with_eval_mean    <- mean(corpus$funs_with_eval)
funs_with_eval_median  <- median(corpus$funs_with_eval)

size_bins       <- c(0L, 1e3L, 5e3L, 1e4L, 5e4L, 1e5L)
size_bins_lasb  <- c("< 1000", "1,000 - 4,999", "5,000 - 9,999", "10,000 - 49,999", "> 50,000")

corpus %>%
  mutate(
    outlier=funs_with_eval > 30 | package_r_code > PACKAGE_SIZE_OUTLIER,
    label=ifelse(rank <= MOST_DOWNLOADED | outlier, package, NA),
    code_size=cut(package_r_code, size_bins, labels=size_bins_lasb, include.lowest=TRUE)
  ) %>%
  ggplot(
    aes(
      x=coverage_expr, 
#      y=funs_with_eval,
      y=evals,
      label=label,
      size=code_size
    )
  ) +
  geom_point(alpha=.4) +
  geom_text(size=3, check_overlap = T, vjust=-1.5, na.rm = TRUE) + 
  geom_hline(aes(yintercept=funs_with_eval_mean), linetype="dashed", color="red", size=0.2) +
  geom_vline(aes(xintercept=coverage_mean), linetype="dashed", color="red", size=0.2) +
  scale_x_continuous(labels = scales::percent) + 
  scale_y_continuous(trans=scales::pseudo_log_trans(), breaks=c(0, 1, 2, 5, 10, 20, 50, 100, 200, 500)) +
  scale_size_discrete(range = c(1, 6)) +
  theme(
    legend.position=c(0.2, 0.85),
    legend.box="horizontal",
    legend.box.background = element_rect(fill="white", size=0.1)
  ) +
  labs(
    x="Code coverage", 
    y="Number of eval call sites", 
    size="Lines of R code",
    linetype=""
  ) +
  guides(
    size=guide_legend(order=1),
    color=FALSE,
    linetype=FALSE
  )

ggsave(CORPUS_PLOT)
```

### List of corpus packages

```{r list of packages, eval=FALSE}
datatable(
  corpus, 
  filter="top", 
  options=list(
    pageLength=50,
    autoWidth=TRUE
  )
)
```

## Kaggle

```{r compute kaggle programs}
kaggle_programs <-
  kaggle %>%
  filter(exitval==0) %>%
  transmute(
    id, 
    code, 
    runtime=as.numeric(runtime, units="seconds")
  )
```

```{r kaggle summary}
kaggle_code     <- kaggle_programs$code
kaggle_runtime  <- kaggle_programs$runtime

overview_table(
  r("kaggle",                 kaggle),
  r("runnable kaggle",        filter(kaggle, runnable)),
  r("finished kaggle",        kaggle_programs),
  r("finished kaggle code",   sum(kaggle_code)),
  r("finished kaggle mean",   mean(kaggle_code)),
  r("finished kaggle median", median(kaggle_code)),
  r("kaggle runtime",         as.duration(sum(kaggle_runtime))),
  r("kaggle mean",            as.duration(mean(kaggle_runtime))),
  r("kaggle median",          as.duration(median(kaggle_runtime)))
)
```

## Combined

```{r compute programs}
programs <- bind_rows(
  corpus_programs %>% mutate(corpus="packages") %>% select(code, runtime, corpus),
  kaggle_programs %>% mutate(corpus="kaggle") %>% select(code, runtime, corpus)
)
```

```{r all summary}
overview_table(
  r("all programs",      nrow(kaggle_programs) + nrow(corpus_programs)),
  r("all runnbale code", sum(kaggle_code) + sum(runnable_code$code)),
  r("all code",          sum(kaggle_code) + sum(corpus$package_code) + sum(runnable_code$code))
)
```

```{r programs-code, eval=F}
programs %>%
  ggplot(aes(y=code, x=corpus, fill=corpus)) + 
  geom_violin() +
  scale_y_log10(labels = scales::comma) +
  scale_x_discrete(labels=c("kaggle"="Kaggle", "packages"="Packages")) +
  scale_fill_viridis(discrete = TRUE) +
  guides(fill = FALSE) +
  labs(y="Code size (log)") +
  theme(
    legend.box="horizontal",
    axis.title.x=element_blank()
  )

ggsave(path(PLOT_DIR, "programs-code.pdf"), width=7, height=5.5)
```

```{r programs-runtime, eval=F}
programs %>%
  ggplot(aes(y=runtime, x=corpus, fill=corpus)) + 
  geom_flat_violin(trim=FALSE, width=.5) +
  geom_dotplot(
    binaxis="y", 
    dotsize=0.1, 
    stackdir="down", 
    binwidth=0.01, 
    position=position_nudge(-0.025)) +
  scale_y_log10(labels = function(x, ...) {
    min <- as.integer(x/60L)
    sec <- as.integer(x%%60L)
    sprintf("%.2d:%.2d", min, sec)
  }) +
  scale_x_discrete(labels=c("kaggle"="Kaggle", "packages"="Packages")) +
  scale_fill_viridis(discrete = TRUE) +
  guides(fill = FALSE) +
  labs(y="Running time [mm:ss] (log)") +
  theme(
    legend.box="horizontal",
    axis.title.x=element_blank()
  )
ggsave(path(PLOT_DIR, "programs-runtime.pdf"), width=3.5, height=5.0)
```