### 2020-08-26 inst/analysis/counts.Rmd
Add potentially useful stuff

---
title: "Empirical Evaluation of eval in R"
output:
  rmdformats::readthedown:
        code_folding: hide
        lightbox: true
        gallery: false
        df_print: paged
        toc_depth: 3
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
params:
    base_dir: /home/aviral/projects/revalstudy2
    kaggle_datafile: "summarized-kaggle.fst"
    core_datafile: "summarized-core.fst"
    packages_datafile: "summarized-packages.fst"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.align = "center",
                      fig.retina = 2,
                      fig.width = 10,
                      cache.lazy = FALSE)

library(tidyverse)
library(stringr)
library(viridis)
library(vctrs)
library(DT)
library(fst)
library(fs)
library(purrr)
library(readr)
library(xtable)

source("inc/paths.R")
source("inc/setup.R")
source("inc/latextags.R")
create_tags(path("../paper/tag", "counts.tex"), prefix="", default=TRUE)
```

```{r eval=TRUE}
run_trace_data <- read_fst(PACKAGE_TRACE_RUNS_FILE)
parallel_kaggle_data <- read_fst(KAGGLE_LOG_FILE)

core_data <- read_fst(params$core_datafile)
packages_data <- 
    read_fst(params$packages_datafile) %>%
    mutate(eval_source_type = "package")
kaggle_data <- 
    read_fst(params$kaggle_datafile)
    #mutate(eval_source_type == "core")

all_data <- bind_rows(core_data, packages_data, kaggle_data)
all_data <-
    all_data %>%
    mutate(eval_call_srcref = 
          coalesce(eval_call_srcref, 
                   str_c("missing", "::", caller_package, "::", caller_function)))

count_summary <-
    all_data %>%
    group_by(eval_source_type, eval_function) %>%
    summarize(call_count = sum(nb_ev_calls),
              site_count = length(unique(eval_call_srcref))) %>%
    ungroup() %>%
    add_row(eval_source_type = "kaggle", eval_function = "eval", call_count = 0, site_count = 0) %>%
    add_row(eval_source_type = "kaggle", eval_function = "eval.parent", call_count = 0, site_count = 0) %>%
    add_row(eval_source_type = "kaggle", eval_function = "evalq", call_count = 0, site_count = 0) %>%
    add_row(eval_source_type = "kaggle", eval_function = "local", call_count = 0, site_count = 0)

eval_source_type_all_summary <-
    count_summary %>%
    group_by(eval_source_type) %>%
    summarize(call_count = sum(call_count),
              site_count = sum(site_count)) %>%
    ungroup() %>%
    mutate(eval_function = "all")

all_eval_function_summary <-
    count_summary %>%
    group_by(eval_function) %>%
    summarize(call_count = sum(call_count),
              site_count = sum(site_count)) %>%
    ungroup() %>%
    mutate(eval_source_type = "all")

all_all_summary <- tibble(
    eval_source_type = "all",
    eval_function = "all",
    call_count = sum(count_summary$call_count),
    site_count = sum(count_summary$site_count)
)

count_summary <-
    bind_rows(count_summary, 
              eval_source_type_all_summary, 
              all_eval_function_summary, 
              all_all_summary)

print(count_summary)

count_summary %>%
pmap_dfr(function(eval_source_type, eval_function, call_count, site_count) {
    eval_function <- str_replace(eval_function, "\\.", " ")
    name <- str_c(eval_source_type, " ", eval_function, " ", "call count")
    r(name, call_count)
    name <- str_c(eval_source_type, " ", eval_function, " ", "site count")
    r(name, site_count)
})
```

```{r run-data, eval=TRUE}
run_eval_proportion <-
    all_data %>%
    group_by(file, eval_source_type) %>%
    summarize(eval_count = sum(nb_ev_calls)) %>%
    ungroup() %>%
    pivot_wider(names_from = "eval_source_type", values_from = "eval_count", values_fill = 0) %>%
    mutate(all = core + package)

total_file_count <- length(unique(run_eval_proportion$file))

no_eval_file_count <-
    run_eval_proportion %>%
    filter(core == 0 & package == 0)

core_eval_file_count <-
    run_eval_proportion %>%
    filter(core != 0 & package == 0)

package_eval_file_count <-
    run_eval_proportion %>%
    filter(core == 0 & package != 0)

all_eval_file_count <-
    run_eval_proportion %>%
    filter(core != 0 & package != 0)

r("total file count", total_file_count)
r("no eval file count", no_eval_file_count)
r("no eval file perc", ratio(no_eval_file_count, total_file_count))
r("core eval file count", core_eval_file_count)
r("core eval file perc", ratio(core_eval_file_count, total_file_count))
r("package eval file count", package_eval_file_count)
r("package eval file perc", ratio(package_eval_file_count, total_file_count))
r("all eval file count", all_eval_file_count)
r("all eval file perc", ratio(all_eval_file_count, total_file_count))

run_eval_proportion <-
    run_eval_proportion %>%
    filter(core != 0 & package != 0) %>%
    mutate(core_perc = round(core * 100 / all)) %>%
    group_by(core_perc) %>%
    summarize(file_count = n()) %>%
    ungroup() %>%
    arrange(desc(core_perc)) %>%
    mutate(file_perc = round(100 * file_count / sum(file_count), 1)) %>%
    mutate(cumsum_file_perc = cumsum(file_perc)) %>%
    mutate(cumsum_file_count = cumsum(file_count))

eighty_core_eval_file_count <-
    run_eval_proportion %>%
    filter(core_perc == 80) %>%
    pull(cumsum_file_count)

r("eighty core eval file count", eighty_core_eval_file_count)
r("eighty core eval file perc", ratio(eighty_core_eval_file_count, all_eval_file_count))

run_eval_proportion
```

```{r package-counts, eval=TRUE}
package_calls <-
    packages_data %>%
    filter(eval_source != "base?") %>%
    group_by(eval_source) %>%
    summarize(site_count = length(unique(eval_call_srcref)), 
              call_count = sum(nb_ev_calls)) %>%
    ungroup() %>%
    arrange(desc(call_count)) %>%
    mutate(call_perc = round(100 * call_count / sum(call_count), 1)) %>%
    mutate(cumsum_call_perc = cumsum(call_perc))

top_ten_package_calls <-
    package_calls %>%
    slice(n = 1:10)

datatable(top_ten_package_calls)

top_ten_package_call_count <- sum(top_ten_package_calls$call_count)
top_ten_package_site_count <- sum(top_ten_package_calls$site_count)
total_package_call_count <- sum(package_calls$call_count)

r("top ten package call count", top_ten_package_call_count)
r("top ten package call perc", ratio(top_ten_package_call_count, total_package_call_count))

r("top ten package site count", top_ten_package_site_count)
r("top ten package site perc", ratio(top_ten_package_site_count, sum(package_calls$site_count)))

top_ten_package_calls %>%
add_column(index = toupper(letters)[1:nrow(top_ten_package_calls)]) %>%
pmap_dfr(function(eval_source, site_count, call_count, call_perc, cumsum_call_perc, index) {
    r(str_c("top ten package name ", index), eval_source)
    r(str_c("top ten package callsite count ", index), site_count)
    r(str_c("top ten package call count ", index), call_count)
    r(str_c("top ten package call perc ", index), ratio(call_count, total_package_call_count))
})
```

```{r callsite-distribution, eval=TRUE}
```{r call-site-table}
eval_callsite_table <-
    package_calls %>%
    mutate(site_count = if_else(site_count <= 10, as.character(site_count), 
                        if_else(site_count <= 50, "11--50",
                        if_else(site_count <= 100, "51--100",
                        if_else(site_count <= 150, "101--150",
                        if_else(site_count <= 200, "151--200",
                        if_else(site_count <= 250, "201--250",
                        if_else(site_count <= 300, "251--300",
                        "> 300")))))))) %>%
    group_by(site_count) %>%
    summarize(package_count = n()) %>%
    ungroup() %>%
    arrange(match(site_count, c(1:10,"11--50", "51--100", "101--150", "151--200", "201--250", "251--300", "> 300")))

datatable(eval_callsite_table)

eval_callsite_table %>%
add_column(index = toupper(letters)[1:nrow(eval_callsite_table)]) %>%
pmap_dfr(function(site_count, package_count, index) {
    r(str_c("site summary site count ", index), site_count)
    r(str_c("site summary package count ", index), package_count)
})
```

```{r core-dependent-own-eval}
all_data <-
    bind_rows(core_data, packages_data, kaggle_data)
    ##filter(is.na(eval_call_srcref))

all_data$package <- map_chr(path_split(all_data$file), function(l) l[[length(l) - 2]])

all_data <-
    all_data %>%
    mutate(core_own_dependent = if_else(eval_source == "core", "core",
                                if_else(eval_source == "base", "core",
                                if_else(package == eval_source, "own",
                                        "dependent"))),
          eval_source)

#all_package_programs <- 
#    run_trace_data %>%
#    filter(Exitval != 0)
#    
#
#all_kaggle_programs <-
#    parallel_kaggle_data %>% 
#    filter(Exitval == 0) %>% 
#    pull(Command) %>% 
#    map_chr(function(str) { 
#        script <- str_split(str, " ")[[1]][4]; 
#        comps <- path_split(script)[[1]]; 
#        str_c(comps[8], "/", comps[9]) 
#    })
#    
#all_files <-
#    rbind(tibble(file = run_trace_data$file, core_own_dependent = "core", corpus = "cran", nb_ev_calls = 0),
#          tibble(file = run_trace_data$file, core_own_dependent = "own", corpus = "cran", nb_ev_calls = 0),
#          tibble(file = run_trace_data$file, core_own_dependent = "dependent", corpus = "cran", nb_ev_calls = 0)) %>%
#    rbind(tibble(file = all_kaggle_programs, core_own_dependent = "core", corpus = "kaggle", nb_ev_calls = 0),
#          tibble(file = all_kaggle_programs, core_own_dependent = "own", corpus = "kaggle", nb_ev_calls = 0),
#          tibble(file = all_kaggle_programs, core_own_dependent = "dependent", corpus = "kaggle", nb_ev_calls = 0))

core_own_dependent <-
    all_data %>%
#        rbind(all_files) %>%
#        select(file, core_own_dependent, corpus, nb_ev_calls) %>%
        group_by(file, core_own_dependent) %>%
        summarize(corpus = first(corpus), eval_calls = sum(nb_ev_calls)) %>%
        ungroup() %>%
        group_by(corpus, core_own_dependent) %>%
        summarize(total_eval_calls = sum(eval_calls),
                  max_eval_calls = max(eval_calls),
                  min_eval_calls = min(eval_calls),
                  avg_eval_calls = mean(eval_calls))

core_own_dependent %>%
pmap_dfr(function(corpus, core_own_dependent, total_eval_calls, max_eval_calls, min_eval_calls, avg_eval_calls) {
    r(str_c(corpus, " ", core_own_dependent, " total eval calls"), total_eval_calls)
    r(str_c(corpus, " ", core_own_dependent, " max eval calls"), max_eval_calls)
    r(str_c(corpus, " ", core_own_dependent, " min eval calls"), min_eval_calls)
    r(str_c(corpus, " ", core_own_dependent, " avg eval calls"), avg_eval_calls)
})

datatable(core_own_dependent)
```