### 2020-08-27 inst/analysis/side-effects.Rmd
Percentages for AST size

---
title: "Side effects"
output: html_document
params:
  base_dir: ../../../
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
for (x in c(
  "dplyr", 
  "DT",
  "fs",
  "fst",
  "ggplot2",
  "ggthemes",
  "lubridate",
  "readr",
  "purrr",
  "stringr",
  "tidyr"
  )) {
  suppressPackageStartupMessages(library(x, character.only=TRUE))
}

knitr::opts_chunk$set(
  echo = TRUE,
  fig.align = "center",
  fig.retina = 2,
  fig.width = 10
)

# TODO: move into the package
source("inc/paths.R")
source("inc/setup.R")
source("inc/latextags.R")

options(repos="https://cran.r-project.org")
theme_set(theme_minimal())
#create_tags(path(TAGS_DIR, "side-effects.tex"), prefix="SE", default=TRUE)
```

```{r}
# FIXME: path!
calls_raw <- read_fst(print(path(params$base_dir, "run", "package-evals-traced.1", "calls.fst")))
calls <- 
  calls_raw %>%
  mutate(package=basename(dirname(dirname(file))))

core_data <- read_fst(EVALS_SUM_CORE_FILE)
package_data <- read_fst(EVALS_SUM_PKGS_FILE)
kaggle_data <- read_fst(EVALS_SUM_KAGGLE_FILE)
program_data <- read_fst(PROGRAM_FILE)

all_data <- rbind(core_data,
                  package_data, 
                  kaggle_data)

head(all_data)

```

## Amount of computation via eval

```{r writes-data}
writes_data <-
    all_data %>%
    select(eval_source_type, eval_source, eval_call_srcref, direct_writes, indirect_writes) %>%
    filter(eval_function == "eval" | eval_function == "evalq") %>%
    mutate(all_writes = direct_writes + indirect_writes, side_effecting = as.logical(all_writes))
# eval_events <-
#  calls %>%
#  filter(!is.na(eval_call_srcref)) %>%
#  filter(eval_function == "eval" | eval_function == "evalq") %>%
#  mutate(all_writes = direct_writes + indirect_writes) %>%
#  select(package, file, eval_call_srcref, eval_call_id, builtin, special, closure, interpreter_eval, c_call, allocation, direct_writes, indirect_writes, all_writes) %>%
#  pivot_longer(cols=c(builtin, special, closure, interpreter_eval, c_call, allocation, direct_writes, indirect_writes, all_writes), names_to="event", values_to="event_count")
```

```{r side-effects-analysis}
eval_calls_by_source_writes <-
    writes_data %>%
    count(eval_source_type, side_effecting, wt = nb_ev_calls, name = call_count)

datatable(eval_calls_by_source_writes)

eval_calls_by_package_writes <-
    writes_data %>%
    filter(eval_source_type != "core") %>%
    count(eval_source, side_effecting, wt = nb_ev_calls, name = call_count) %>%
    arrange(desc(call_count))

datatable(eval_calls_by_package_writes)

eval_sites_by_source_writes <-
    writes_data %>%
    group_by(eval_source_type, eval_call_srcref) %>%
    summarize(side_effecting = any(side_effecting)) %>%
    ungroup() %>%
    count(eval_source_type, side_effecting, name = site_count)

datatable(eval_sites_by_source_writes)

eval_sites_by_package_writes <-
    writes_data %>%
    filter(eval_source_type != "core") %>%
    group_by(eval_source, eval_call_srcref) %>%
    summarize(side_effecting = any(side_effecting)) %>%
    ungroup() %>%
    count(eval_source, side_effecting, name = site_count) %>%
    group_by(eval_source, side_effecting) %>%
    mutate(site_perc = round(100 * site_count / sum(site_count), 1)) %>%
    ungroup() %>%
    arrange(side_effecting, desc(site_perc))

datatable(eval_sites_by_package_writes)

```

Each entry in `call.fst` contains `direct_writes` and `indirect_writes` (both `0` for `eval.parent` and `local`). 
`direct_writes` are the side-effects directly caused by this `eval`. 
`indirect_writes` are the side-effects caused by all children evals of this eval.

For graphing, you may want to sum them up for each eval.
I compute them separately because it is more informative (should we need this information for debugging or qualitative insights).
Now, for normalization you have two quantities: all assignments done by the application and all assignments done by all evals (non-local or local).
all application assignments can be found in `direct_writes` field of `program.fst` and all `eval_assignments` can be found in `indirect_writes` field of `program.fst`.
You can also find all non-local assignments by simply summing up the `direct_writes` field of all eval calls.

```{r}

```