### 2020-08-27 inst/analysis/analysis_stable.Rmd
Environments

---
title: "Empirical Evaluation of eval in R"
output:
  rmdformats::readthedown:
        code_folding: hide
        lightbox: true
        gallery: false
        df_print: paged
        toc_depth: 3
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
params:
  base_dir: /var/lib/R/project-evalR

# This makes it possible to generate the html file in the html subfolder
# knit: (function(inputFile, encoding) {
#   rmarkdown::render(inputFile, encoding = encoding, output_dir = "html") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.align = "center",
                      fig.retina = 2,
                      fig.width = 10,
                      cache.lazy = FALSE)

now <- Sys.time()

library(tidyverse)
library(stringr)
library(viridis)
library(vctrs)
library(DT)
library(fst)
library(fs)
library(ggthemes)

source("insights.R", local = knitr::knit_global())

source("inc/paths.R")
source("inc/latextags.R")
source("inc/setup.R")
theme_set(theme_minimal())
theme_update(text = element_text(size = 20))

create_tags(path(TAGS_DIR, "analysisStable.tex"), prefix="", default=TRUE)

# Temp 
# RUN_DIR   <- path(params$base_dir, "run")
# RUN_TRACE_DIR  <- path(RUN_DIR, "package-evals-traced")
# EVALS_RAW_FILE           <- path(RUN_TRACE_DIR, "raws.fst")
# EVALS_SUM_CORE_FILE      <- path(RUN_TRACE_DIR, "summarized-core.fst")
# EVALS_SUM_PKGS_FILE      <- path(RUN_TRACE_DIR, "summarized-packages.fst")
# EVALS_SUM_KAGGLE_FILE    <- path(RUN_TRACE_DIR, "summarized-kaggle.fst")
# EVALS_SUM_EXTERNALS_FILE <- path(RUN_TRACE_DIR, "summarized-externals.fst")
packages_filepath <- EVALS_SUM_PKGS_FILE
core_filepath <- EVALS_SUM_CORE_FILE
```

# Data

The analysis is being performed on dataset read from files `r packages_filepath` and `core_filepath`.

```{r file-info-packages, echo=FALSE}
file_info(packages_filepath)
```

```{r file-info-core, echo=FALSE}
file_info(core_filepath)
```

```{r load-data, echo=FALSE}
eval_calls_packages <-
    read_fst(packages_filepath) %>%
    as_tibble()  %>% 
    mutate(dataset = "packages")

eval_call_core <- 
  read_fst(core_filepath) %>%
  as_tibble() %>%
  mutate(dataset = "core")

eval_calls_raw <- bind_rows(eval_calls_packages, eval_call_core)
```

We remove a few columns we do not need here and deduplicate.
```{r deduplicate}
# eval_calls <- eval_calls_raw %>% count(across(-c(builtin:indirect_allocation)), name = "nb_ev_calls2") %>%
#     mutate(nb_ev_calls = nb_ev_calls * nb_ev_calls2) %>% 
#     select(-nb_ev_calls2)

#
eval_calls <- eval_calls_raw
```

```{r document}
# A flexible way to document the dataset: it is quick and easy to document a new column
c_names <- names(eval_calls)
column_names <- setNames(data.frame(matrix(ncol = length(c_names), nrow = 0)), c_names)

column_names <- column_names %>% add_row()

column_names["file"] <- "File where the `eval` is executed"
column_names["eval_call_id"] <- "Unique id of a call" 
column_names["expr_parsed_expression"] <- "Contains the expression with a call to `parse` if there is one in the `expr` argument somewhere in the call stack."
column_names["caller_function"] <- "Function in which the `eval` is called."
column_names["caller_package"] <- "Package in which the `eval` is called."
column_names["eval_function"] <- "One of `eval`, `evalq`, `eval.parent`, and `local`."
column_names["eval_call_expression"] <- "The full expression with `eval` that has been traced."
column_names["eval_call_srcref"] <- "File name and line where the `eval` call is. Might ne NA, especially for calls originating from package base."
column_names["caller_expression"] <- "Expression in which the `eval` is called."
column_names["caller_srcref"] <- "File name and line where the caller of the `eval` is located. Might be NA (especially for callers in base)."
# The following ones were in package-evals-traced-corpus.1/calls.fst
#column_names["caller_srcref2"] <- "File name and line for some calls for which the caller seemed to be itself (mainly Rcpp and C calls)."
#column_names["caller_expression_2"] <- "Caller expression when the caller seems to call itself (see `caller_srcref2`)."
#column_names["caller_expression_2"] <- "Same as `caller_expression_2` but uses `format` to print, not `expr_to_string`."
column_names["caller_stack_expression"] <-"When `eval_function` == `caller_function`, climb up the stack to find the first different caller."
column_names["caller_stack_expression_raw"] <- "See `caller_stack_expression`. Does not use `expr_to_str` but `format`."
column_names["caller_stack_expression_srcref"] <- "srcref for `caller_stack_expression`."
# To continue

# Pivoting

column_names <- column_names %>% rownames_to_column() %>% 
  pivot_longer(-rowname) %>%
  pivot_wider(names_from=rowname, values_from=value)

column_names <- column_names %>%  rename("Description" = `1`) %>% filter(!is.na(Description))

```

```{r unique_calls}
nb_unique_calls <- nrow(eval_calls)

r("unique calls", nb_unique_calls)
```

There are `r nb_unique_calls` unique calls;

The columns of the dataset are the following ones:
```{r column_names}
column_names %>% knitr::kable()
```

# Corpus

TODO

# Research Questions

## RQ1: How many evals are there?

There are four kinds of `eval` in the `base` package. Their definitions are shown below.

```{r, eval-definition, eval=FALSE}
base::eval <-
    function(expr, 
             envir = parent.frame(), 
             enclos = if (is.list(envir)||is.pairlist(envir)) parent.frame() else baseenv())
        .Internal(eval(expr, envir, enclos))

base::evalq <-
    function (expr, 
              envir = parent.frame(), 
              enclos = if (is.list(envir) || is.pairlist(envir)) parent.frame() else baseenv())
        .Internal(eval(substitute(expr), envir, enclos))

base::eval.parent <-
    function (expr, n = 1) {
        p <- parent.frame(n + 1)
        eval(expr, p)
    }

base::local <-
    function (expr, envir = new.env())
        eval.parent(substitute(eval(quote(expr), envir)))
```

```{r eval-count-table}
callsites_with_srcref <-
    eval_calls %>% 
    select(eval_call_srcref) %>% 
    n_distinct(na.rm = TRUE)

callsites_without_srcref <-
    eval_calls %>% 
    filter(is.na(eval_call_srcref)) %>% 
    select(caller_package, caller_function) %>% 
    n_distinct()

eval_calls <-
    eval_calls %>%
    mutate(eval_call_srcref = coalesce(eval_call_srcref, 
                                       str_c("missing", "::", caller_package, "::", caller_function)))
eval_count_table <-
    eval_calls %>%
    group_by(eval_function, eval_call_srcref) %>%
    summarize(count = sum(nb_ev_calls)) %>%
    summarize(call_count = sum(count), callsite_count = n())

nb_eval_calls <- total_call_count <- sum(eval_count_table$call_count)

nb_call_sites <- total_callsite_count <- sum(eval_count_table$callsite_count)

eval_count_table <-
    eval_count_table %>%
    mutate(call_proportion = 100 * call_count / total_call_count,
           callsite_proportion = 100 * callsite_count / total_callsite_count) %>%
    arrange(desc(call_count))

eval_call_proportion <- 
    eval_count_table %>%
    filter(eval_function == "eval") %>%
    pull(call_proportion)
```

There are `r NA` R function calls with `r total_call_count` calls to the four kinds of eval which accounts for `r NA`% of all R function calls.
The table below summarizes the proportion of calls to the different kinds of evals.

```{r }
eval_count_table %>%
select(eval_function, call_count, call_proportion) %>% 
datatable()
```

We observe that `r eval_call_proportion`% of the calls are made to the `eval` function.

These `eval` calls originate from `r total_callsite_count` call sites. For the number of call sites, we use line number information attached to the AST by the parser. When it is not available (such as in *base*), we use a combination of `caller_package` and `caller_function`. We assume that:

  - the combination `caller_package` and `caller_function` is unique
  - there is only one call site per such `caller_function`
This assumption results in a lower bound for the number of unique callsites for eval.

```{r call-site-table}
eval_callsite_table <-
    eval_calls %>%
    count(eval_function, caller_package, eval_call_srcref, name = "call_count") %>%
    group_by(eval_function, caller_package) %>%
    summarize(call_count = sum(call_count), callsite_count = n()) %>%
    ungroup()

eval_callsite_table %>%
datatable()
```

```{r call-site-distribution}
package_callsite_distribution <-
    eval_callsite_table %>%
    mutate(callsite_count = if_else(callsite_count <= 5, as.character(callsite_count), "> 5")) %>%
    group_by(eval_function, callsite_count) %>%
    summarize(package_count = n())

package_callsite_distribution %>%
datatable()

p <-
    package_callsite_distribution %>%
    ggplot(aes(callsite_count, package_count)) +
    geom_col() +
    facet_wrap(vars(eval_function))
p
```

To asses how precise the call site estimate is for calls that do not have `srcref`, we compare it on *base* with the actual count given by `grep`.

```{r eval_call_sites_base}
nb_call_site_base <- eval_calls %>% filter(caller_package == "base") %>% select(caller_function) %>% n_distinct(na.rm = TRUE)
```
The estimate for *base* is `r nb_call_site_base` call sites, while `grep` yields 40 `eval` call sites.

For the other packages, ranked by call sites:
```{r eval_call_sites_per_package}
eval_calls %>% filter(caller_package != "base") %>% group_by(caller_package) %>% summarize(nb_call_sites =  n_distinct(eval_call_srcref, na.rm = TRUE)) %>% arrange(desc(nb_call_sites))
```

We probably forget the generated ones here, but should we count them in the _call sites_?

What are the packages with the highest number of `eval` calls?

```{r nb_eval_calls_per_package}
nb_eval_calls_per_package <- eval_calls %>% group_by(caller_package) %>% summarise(nb_eval_calls_package = sum(nb_ev_calls), per_cent = nb_eval_calls_package / nb_eval_calls * 100) %>% arrange(desc(nb_eval_calls_package))
```

*base* has the highest number of eval calls, i.e. `r nb_eval_calls_per_package[1, "per_cent"][[1]]`% of them.

```{r}
nb_eval_calls_per_package  %>% datatable()
```

What are the most used caller functions?

```{r caller_functions_used}
caller_function_ranked <- eval_calls %>% count(caller_function, wt = nb_ev_calls, sort = TRUE) %>% mutate(proportion = n / sum(n))

caller_function_ranked %>% datatable()

```

The most used function is `r caller_function_ranked[1, "caller_function"][[1]]`, as much as `r 100 * caller_function_ranked[1, "proportion"][[1]]`%.

```{r match_arg}
match_arg_r <- caller_function_ranked %>% filter(caller_function == "match.arg")

r("match arg number", match_arg_r[1, "n"][[1]])
r("match arg percent", percent(match_arg_r[1, "proportion"][[1]]))
```

### Excluding *base* from the data set

When using `grep` , we find 30 call sites that use `eval` in *base*.
We tried to replace the 30 `eval` call sites in *base* and could do it in  50% of the cases.

We reckon that `eval` in *base* should be considered apart from the other `eval` calls.

`eval` in *base* does not consume user input, i.e. input coming from `parse`, except in 5 legitimate cases, such as the `sys.source` function that is used to `eval` a file.

| Function name              | File           | Line | Parse ? | Comments                                                                           |
|----------------------------|----------------|------|---------|------------------------------------------------------------------------------------|
| autoload                   | autoload       | 25   |         |                                                                                    |
| autoloader                 | autoload       | 43   |         | To execute a call to library                                                       |
| autoloader                 | autoload       | 49   |         |                                                                                    |
| bquote                     | backquote      | 33   |         | In subfunction unquote. To eval what is in .() . Metaprogramming                   |
| bquote                     | backquote      | 56   |         | idem (in unquote.list, defined inside bquote)                                      |
| by.default                 | by             | 35   |         | eval(substitute pattern?                                                           |
| by.data.frame              | by             | 50   |         | idem                                                                               |
| invokeRestartInteractively | conditions     | 245  | yes     | prompt                                                                             |
| invokeRestartInteractively | conditions     | 249  | yes     | prompt                                                                             |
| Ops.data.frame             | dataframe      | 1678 |         | for arithmetic operators on data frames ; do.call could be probably used here also |
| dget                       | dput           | 32   | yes     | from file                                                                          |
| eval                       | eval           | 26   |         | Call Internal eval                                                                 |
| eval.parent                | eval           | 30   |         |                                                                                    |
| evalq                      | eval           | 36   |         | evalq is basically eval(substitute                                                 |
| local                      | eval           | 49   |         |                                                                                    |
| with.default               | eval           | 57   |         |                                                                                    |
| within.data.frame          | eval           | 63   |         |                                                                                    |
| within.list                | eval           | 77   |         | evalq                                                                              |
| replicate                  | replicate      | 21   |         |                                                                                    |
| subset.data.frame          | frametools     | 25   |         |                                                                                    |
| subset.data.frame          | frametools     | 34   |         |                                                                                    |
| subset.matrix              | frametools     | 56   |         |                                                                                    |
| transform.data.frame       | frametools     | 69   |         |                                                                                    |
| match.arg                  | match          | 45   |         |                                                                                    |
| char.expand                | match          | 79   |         |                                                                                    |
| max.col                    | max            | 22   |         | To access formals                                                                  |
| parseNamespaceFile         | namespace      | 1315 | yes     | To parse directives                                                                |
| idem                       | namespace      | 1319 | yes     |                                                                                    |
| idem                       | namespace      | 1442 | yes     | evaluate a call from a parse                                                       |
| source                     | source         | 123  | yes     | Use the internal parse function (here it is a comment)                             |
| source                     | source         | 170  | yes     | comment                                                                            |
| source                     | source         | 219  | yes     | the actual eval                                                                    |
| sys.source                 | source         | 277  | yes     | the non-internal parse is used                                                     |
| stopifnot                  | stop           | 54   |         | change of caller pattern                                                           |
| as.data.frame.table        | table          | 244  |         | Build a call?                                                                      |
| match.fun                  | match.fun      | 28   |         | eval.parent                                                                        |
| trace                      | methodsSupport | 38   |         | eval.parent                                                                        |
| untrace                    | methodsSupport | 53   |         | eval.parent                                                                        |
| .doTrace                   | methodsSupport | 78   |         | eval.parent                                                                        |

```{r parse_in_base}
parsed_base <- eval_calls %>% filter(!is.na(expr_parsed_expression), eval_source == "base")

nb_call_site_parse_base <-  nb_eval_call_sites(parsed_base)
```

There are `r count(parsed_base, wt = nb_ev_calls)` `eval` calls with `parse` in *base*, which is `r count(parsed_base, wt = nb_ev_calls) / count(filter(eval_calls, eval_source == "base"), wt=nb_ev_calls)`%.

## Size of expressions as arguments to `eval`

### String size

```{r}
eval_calls %>% ggplot() +
  geom_histogram(aes(x = expr_expression_length, weight = nb_ev_calls), na.rm = TRUE, binwidth=1) +
  scale_x_log10() +  scale_y_sqrt()
```

If we remove the outliers:
```{r}
eval_calls %>% filter(expr_expression_length < 25000) %>% ggplot() +
  geom_histogram(aes(x = expr_expression_length, weight = nb_ev_calls), na.rm = TRUE, binwidth=2)  + scale_x_sqrt() +  scale_y_sqrt() 
```

What about the code size of expressions that seem to be generated?
An approximation of *generated* will be no `srcref` and not in `base`.

```{r code_size_srcref}
eval_calls %>% filter(is.na(eval_call_srcref), caller_package != "base") %>% ggplot() +
  geom_histogram(aes(x = expr_expression_length, weight = nb_ev_calls), na.rm = TRUE, binwidth=1) + scale_x_sqrt() + scale_y_sqrt()
```

And the average size per package:
```{r average_size_package}
eval_calls %>% group_by(caller_package) %>% summarize(average_expr_size = mean(expr_expression_length, na.rm = TRUE)) %>% arrange(desc(average_expr_size))
```

If we look at the maxima, the largest ones are in *base*:
```{r max_size_package}
eval_calls %>% group_by(caller_package) %>% summarize(max_expr_size = max(expr_expression_length, na.rm = TRUE)) %>% arrange(desc(max_expr_size))
```

### AST size

```{r}
ast_sizes <- eval_calls %>% slice_sample(n=10000) %>% mutate(expr_resolved_nodes = map_int(expr_resolved, expr_size_str))
# Now precomputed during the tracing
#ast_sizes <- eval_calls
```

```{r}
# ast_sizes %>% ggplot(aes(x = expr_resolved_nodes, weight = nb_ev_calls, group = dataset)) +
#   geom_histogram(aes(y = stat(width*density)), na.rm = TRUE, binwidth = 0.1) +
#   scale_x_log10() +  scale_y_sqrt(labels=scales::percent) +
#   labs(x = "AST size", y = "relative frequencies")

# ast_sizes %>% 
#   ggplot(aes(x = dataset, y = expr_resolved_nodes, fill = dataset, weight = nb_ev_calls)) +
#   geom_flat_violin(trim=FALSE, width=.5) +
#   geom_dotplot(
#     binaxis="y", 
#     dotsize=0.01, 
#     stackdir="down", 
#     binwidth=0.01, 
#     position=position_nudge(-0.025)) +
#   scale_y_log10(labels = scales::comma) +
#   labs(y="AST size (log)") +
#   theme(
#     legend.box="horizontal",
#     axis.title.x=element_blank()
#   )

ast_sizes %>%
  ggplot(aes(x = dataset, y = expr_resolved_nodes, fill = dataset, weight = nb_ev_calls)) +
  geom_violin(trim = FALSE, width=.5)+
  geom_boxplot(width=.05) +
  scale_y_log10(labels = scales::comma) +
  labs(y="AST size (log)") +
  theme(
    legend.box="horizontal",
    axis.title.x=element_blank()
  )

ggsave(path(PLOT_DIR, "ast_sizes.pdf"))

ast_sizes_prop <- ast_sizes %>% count(expr_resolved_nodes, wt = nb_ev_calls) %>%
                      mutate(proportion = n / sum(n)) %>% 
                      arrange(expr_resolved_nodes) %>% 
                      mutate(cumulative = cumsum(proportion))

ast_size_one <- ast_sizes %>% filter(expr_resolved_nodes == 1) 
nb_ast_size_one <- count(ast_size_one, wt = nb_ev_calls)$n
nb_ast_size_one_call_sites <- nb_eval_call_sites(ast_size_one)
r("nb ast one percent", ratio(nb_ast_size_one, sum(ast_sizes_prop$n)))
r("nb ast one call site percent", ratio(nb_ast_size_one_call_sites, nb_eval_call_sites(ast_sizes)))

nb_nodes_many <- ast_sizes_prop %>% filter(cumulative >= 0.95) %>% .[1,]

r("ast size nine five", nb_nodes_many$expr_resolved_nodes)
r("ast size nine five exact percent", percent(nb_nodes_many$cumulative))

```

## Amount of computations performed in `eval`

## Aliases of `eval`

# A taxonomy of `eval`

## The operation mix

## Scope (environments)

One of the main differences of the R \eval with the javascript one is how it interacts with \emph{scope}, or environments in R. In Javascript, \c{eval} can access the local scope and the global scope. In R, \c{eval} has an argument \c{envir} to indicate in which environment its first argument, after being resolved, must be evaluated.

By default, it is equal to \c{parent.frame()}, which is the parent environment of the \c{eval} call.

```{r default_environment}
default_envir <- eval_calls %>% filter(is.na(envir_expression)) 
nb_default_envir <- default_envir %>% count(wt = nb_ev_calls)
nb_call_sites_default_envir <- nb_eval_call_sites(default_envir)

r("default envir expr", nb_default_envir$n)
r("default envir expr percent", ratio(nb_default_envir$n, nb_eval_calls))
r("default envir  call sites percent", ratio(nb_call_sites_default_envir, nb_call_sites))
```

`r ratio(nb_default_envir$n, nb_eval_calls)`% of the calls use the default environment, which corresponds to `r ratio(nb_call_sites_default_envir, nb_call_sites)`% of the call sites.

We also classify environments in the following way:

- `base`: environment of a primitive function of *base* package
- `empty`: empty environment
- `callee`: same environment (it seems it only happens with `local` )
- `global`: ` .GlobalEnv` 
- `package:package_name`: package environment
- `loop`: there was a loop in the frame

```{r environment_hierarchy}
# classified_envir <- eval_calls %>% filter(!is.na(envir_expression)) %>%
#       mutate(env_class = pmap(list(environment_class, envir_type, envir_expression), extract_envir)) %>%
#       hoist(env_class, read_env = 1, write_env = 2, .remove=FALSE) #Quicker when not removing

classified_envir <- eval_calls %>% filter(!is.na(envir_expression)) %>% 
  mutate(env_class = pmap_chr(list(environment_class, envir_type, envir_expression), simplify_envir)) %>% 
  mutate(env_class_read = map_chr(env_class, ~ str_match(., "([^\\+]*)\\+?")[[2]]))
```

```{r envirall}
classified_envir %>% ggplot() +
  geom_bar(aes(x =fct_rev(fct_infreq(env_class)), weight = nb_ev_calls )) +
  coord_flip() +
  scale_y_log10() +
  labs(x = "Environments") + 
  facet_wrap(~dataset)

ggsave(path(PLOT_DIR, "environments.pdf"))
```

```{r envir_read}
classified_envir %>% ggplot() +
  geom_bar(aes(x =fct_rev(fct_infreq(env_class_read)), weight = nb_ev_calls )) +
  coord_flip() +
  scale_y_log10() +
  labs(x = "Environments") + 
  facet_wrap(~dataset)

ggsave(path(PLOT_DIR, "environments-read.pdf"))
```

## Patterns

We look at the first argument of `eval`, the `expr` argument, after it is resolved. It is the expression to be evaluated. It can be of different types, mainly usual numeric types, vectors, strings, and language types: calls, symbols, promise, bytecode and expressions.

`bytecode` refers to bytecode objects, that are generated by the `compile`function:

```{r}
typeof(compiler::compile(1+1))
```

Type `expression` is a sequence of calls and symbols.

```{r types_resolved}
per_expr_resolved_type <- eval_calls %>% count(expr_resolved_type, wt = nb_ev_calls)
```

```{r}
 per_expr_resolved_type %>% mutate(per_cent = n / nb_eval_calls * 100 ) %>% arrange(desc(n)) %>% knitr::kable()
```

```{r}
per_expr_resolved_type %>% ggplot() +
  geom_col(aes(x = fct_reorder(expr_resolved_type, n), y = n)) +
  scale_y_sqrt() + 
  coord_flip() +
  labs(title = "Calls per type", y = "n", x = "type" ) 
```

There is a much larger variety of types for the resolved expressions.

```{r not_unchanged_resolved_eval}
changed_resolved_evals <- eval_calls %>% mutate(changed =  fct_collapse(expr_resolved_type, lang_types = c("LANGSXP", "EXPRSXP", "SYMSXP", "PROMSXP", "BCODESXP"), other_level = "other_types")) %>% count(changed, wt = nb_eval_calls) %>% mutate(per_cent = n / nb_eval_calls * 100 ) 
```

We can remove `eval` and leave there the expression in `r changed_resolved_evals[2, "per_cent"][[1]]`% of the eval calls (i.e. `r changed_resolved_evals[2, "n"][[1]]`).

We can do the same analysis as for the non-resolved expressions:

```{r expr_resolved__call_arg}
expr_resolved_call_arg <- eval_calls %>% filter(expr_resolved_type == "LANGSXP")
```

The following top level functions are used:

```{r top_level_function_resolved_expr}
expr_resolved_call_arg %>% count(expr_resolved_function, wt=nb_ev_calls) %>% arrange(desc(n))
```

We can distinguish a few groups:

- assignments
- reading (in a data frame, a vector)
- parsing
- defusing (`quote`, `substitute` and so on)
- conversion or creation of new values
- building calls and calling
- working with environments
- class assignments, reads and definitions
- building logical values
- formula
- meta programming on functions (getting access to the body of the function, to the call stack)
- 'NA' corresponds to no function calls (or that it was not parsed successfully)
- anonymous functions

```{r group_expr_resolved_functions}
group_expr_resolved_functions <- expr_resolved_call_arg %>%
                            mutate(function_expr2 = groupify_function(expr_resolved_function)) %>%
                            mutate(group_function_expr = fct_collapse(function_expr2, 
                              assignment = c("<-", "[[<-", "[<-", "$<-", "<<-", "=", "@<-"),
                              slot_access = c("[", "[[", "$", "slot", "@", "::", ":::"),
                              parsing = c("parse", "str2lang", "str2expression", "parse_only", "base::parse"),
                              defusing = c("quote", "substitute", "bquote", "expr", "quo"),
                              expressions = c("as.list", "expression", "as.name", "as.expression", "list", "c", "as.data.frame", "as.character", "{", "modify_lang"),
                              calling = c("call", "as.call", "do.call"),
                              environment = c("environment", "as.environment"),
                              logical = c("!", "&", "|", "&&", "||"),
                              relation = c("<", ">", "is.element", "<=", "=>", "!="),
                              choice = c("if"),
                              anonymous = c("anonymous"),
                              primitive = c("primitive"),
                              function_meta = c("sys.calls", "function", "body", "missing"),
                              arithmetic = c("+", "-", "*", "/", "%%", "^"),
                              formula = c("~", "deriv", "D", "model.frame", "rootftfunc", "contest.lmerModLmerTest", "JacFunc", "gamlss.family", "contest1D.lmerModLmerTest", "contestMD.lmerModLmerTest", "reexpr"),                           
                              other_level = "other" ))
```

```{r}
group_expr_resolved_functions %>% count(group_function_expr, wt = nb_ev_calls) %>% arrange(desc(n)) %>% knitr::kable()
```

```{r}
group_expr_resolved_functions %>% ggplot() +
  geom_bar(aes(x = fct_rev(fct_infreq(group_function_expr)), weight = nb_ev_calls)) +
  scale_y_sqrt() +
  coord_flip() +
  labs(title = "Category of calls", y = "n", x = "category" ) 
```

```{r }
group_expr_resolved_functions %>% ggplot() +
  geom_bar(aes(x = fct_rev(fct_infreq(group_function_expr)), weight = nb_ev_calls)) +
  scale_y_sqrt() +
  coord_flip() +
  labs(title = "Category of calls", y = "n", x = "category" ) +
  facet_wrap(vars(eval_function))
```

We also look at the arity of the top-level function:

```{r arity_expr_resolved}
expr_resolved_call_arg %>% count(expr_expression_args_num, sort = TRUE) %>% knitr::kable()
```

```{r arity_expr_resolved_plot}
expr_resolved_call_arg %>% ggplot() + 
  geom_bar(aes(x = expr_resolved_args_num, weight = nb_ev_calls)) +
  scale_y_sqrt() +
  coord_flip() + 
  labs(title = "Arity of calls", x = "arity", y = "n")

```

```{r arity_expr_resolved_plot_per_eval}
expr_resolved_call_arg %>% ggplot() + 
  geom_bar(aes(x = expr_resolved_args_num, weight = nb_ev_calls)) + 
  scale_y_sqrt() +
  coord_flip() + 
  labs(title = "Arity of calls", x = "arity", y = "n") +
  facet_wrap(vars(eval_function))
```

Another interesting figure is the degreee of polymorphism of the `expr` argument: how many different types (for `expr_resolved`) are there for one call sites?

```{r polymorphism_callsite}
expr_resolved_polymorphism <- eval_calls %>% group_by(eval_call_srcref) %>%
        summarise(polym_degree = vec_unique_count(expr_resolved_type))

expr_resolved_monomorphic <- expr_resolved_polymorphism %>% filter(polym_degree == 1)
nb_monomorphic <- expr_resolved_monomorphic %>% nrow()

r("nb monomorphic", nb_monomorphic)
r("percent monomorphic", ratio(nb_monomorphic, nb_call_sites))
```

```{r polymorphism}
expr_resolved_polymorphism %>% ggplot() +
  geom_bar(aes(x = polym_degree)) +
  scale_x_continuous(breaks=c(0, 1, 5, 10)) +
  labs(title = "Degree of polymorphism", x = "degree", y = "n")

ggsave(path(PLOT_DIR, "polymorphism.pdf"))
```

`r nb_monomorphic` call sites are not polymorphic, which represents `r nb_monomorphic / nb_call_sites * 100`% of the call sites.

From the monomorphic functions, we can deduce how many are _replaceable_:

```{r}
expr_resolved_replaceable <- expr_resolved_monomorphic %>% left_join(eval_calls) %>% mutate(can_replace = map_lgl(expr_resolved,  is_replaceable_str))
```

## Provenance

Some `eval` result from parsing text (directly or through a file) with `parse`.

```{r eval_from_parse}
parsed_evals <- eval_calls %>% filter(!is.na(expr_parsed_expression))
nb_parsed_evals <- parsed_evals %>% count(wt = nb_ev_calls)

r("percent parsed evals",  ratio(nb_parsed_evals$n, nb_eval_calls))
```

There are `r nb_parsed_evals` such calls, `r nb_parsed_evals / nb_eval_calls * 100`%.

```{r, echo = FALSE}
nb_parsed_call_sites <- nb_eval_call_sites(parsed_evals)

r("percent parsed call sites", ratio(nb_parsed_call_sites, nb_call_sites))
```

In terms of call sites, there are `r nb_parsed_call_sites` i.e. `r nb_parsed_call_sites / nb_call_sites * 100`%.

  We can differentiate further the provenance, i.e. whether the evaluated string comes from `file` or from `text`.

```{r}
if("parse_args_file" %in% names(parsed_evals))
{
  parsed_files <- parsed_evals %>% filter(!is.na(parse_args_file))
  nb_file <- parsed_files %>% count(wt = nb_ev_calls)
  nb_file_call_sites <- nb_eval_call_sites(parsed_files)
} else
{
  nb_file <- 0
  nb_file_call_sites <- 0
}

```

```{r}
parsed_texts <- parsed_evals %>% filter(!is.na(parse_args_text) | !is.na(parse_args_s))
nb_text <-  parsed_texts %>% count(wt = nb_ev_calls)
nb_text_call_sites <- nb_eval_call_sites(parsed_texts)
```

There are `r nb_file` calls with the `file` argument, and `r nb_text` with the `text` argument, i.e. `r  nb_file / nb_parsed_evals * 100`%, and `r nb_text / nb_parsed_evals *100`% respectively.

There are `r nb_file_call_sites` call sites with the `file` argument, and `r nb_text_call_sites` with the `text` argument, i.e. `r  nb_file_call_sites / nb_parsed_call_sites * 100`%, and `r nb_text_call_sites / nb_parsed_call_sites *100`% respectively.

We also look to the occurrences of `paste` (and `parse0`), `glue` and `str_c` in the `parse` argument. It is an indication that meta-pogramming happens.

```{r parse_meta_prog}
meta_parsed <- parsed_texts %>% filter(str_detect(expr_parsed_expression, "paste|str_c|glue"))

nb_meta_parsed <- meta_parsed %>% count(wt = nb_ev_calls)
nb_meta_parsed_call_sites <- nb_eval_call_sites(meta_parsed)

r("paste parsed evals", ratio(nb_meta_parsed$n,  nb_parsed_evals$n))
r("paste parsed call site", ratio(nb_meta_parsed_call_sites,  nb_parsed_call_sites))
```

We can look at the most common meta-programming examples:

```{r parse_meta_prog_ex}
meta_parsed %>% count(expr_parsed_expression, wt = nb_ev_calls, sort = TRUE)
```

## Consistence

# Other faces of `eval`

# Case studies

```{r}
duration <- difftime(Sys.time(), now)
```

Notebook execution was `r duration` `r units(duration)`.