### 2020-08-27 inst/analysis/counts.Rmd

---
title: "Empirical Evaluation of eval in R"
output: html_document
params:
  base_dir: ../../../
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.align = "center",
                      fig.retina = 2,
                      fig.width = 10,
                      cache.lazy = FALSE)

library(tidyverse)
library(stringr)
library(viridis)
library(vctrs)
library(DT)
library(fst)
library(fs)
library(purrr)
library(readr)
library(xtable)

source("inc/paths.R")
source("inc/setup.R")
source("inc/latextags.R")

create_tags(path(TAGS_DIR, "counts.tex"), prefix="", default=TRUE)
```

```{r load programs}
cran_programs <- read_fst(CRAN_PROGRAMS_FILE) %>% as_tibble() %>% filter(success)
kaggle_programs <- read_fst(KAGGLE_PROGRAMS_FILE) %>% as_tibble() %>% filter(success)
programs <- bind_rows(
  select(cran_programs, file, corpus),
  select(kaggle_programs, file, corpus)
)
```

```{r eval=TRUE}
core_data <- 
  read_fst(EVALS_SUM_CORE_FILE) %>%
  as_tibble() %>%
  semi_join(programs, by="file")

packages_data <- 
  read_fst(EVALS_SUM_PKGS_FILE) %>%
  as_tibble() %>%
  semi_join(programs, by="file")

# kaggle should be always empty
kaggle_data <- 
  read_fst(EVALS_SUM_KAGGLE_FILE) %>%
  as_tibble() %>%
  semi_join(programs, by="file")

all_data <- 
  bind_rows(core_data, packages_data, kaggle_data)

all_data <-
    all_data %>%
    mutate(eval_call_srcref = 
          coalesce(eval_call_srcref, 
                   str_c("missing", "::", caller_package, "::", caller_function)))

count_summary <-
    all_data %>%
    group_by(eval_source_type, eval_function) %>%
    summarize(call_count = sum(nb_ev_calls),
              site_count = length(unique(eval_call_srcref))) %>%
    ungroup() %>%
    add_row(eval_source_type = "kaggle", eval_function = "eval", call_count = 0, site_count = 0) %>%
    add_row(eval_source_type = "kaggle", eval_function = "eval.parent", call_count = 0, site_count = 0) %>%
    add_row(eval_source_type = "kaggle", eval_function = "evalq", call_count = 0, site_count = 0) %>%
    add_row(eval_source_type = "kaggle", eval_function = "local", call_count = 0, site_count = 0)

eval_source_type_all_summary <-
    count_summary %>%
    group_by(eval_source_type) %>%
    summarize(call_count = sum(call_count),
              site_count = sum(site_count)) %>%
    ungroup() %>%
    mutate(eval_function = "all")

all_eval_function_summary <-
    count_summary %>%
    group_by(eval_function) %>%
    summarize(call_count = sum(call_count),
              site_count = sum(site_count)) %>%
    ungroup() %>%
    mutate(eval_source_type = "all")

all_all_summary <- tibble(
    eval_source_type = "all",
    eval_function = "all",
    call_count = sum(count_summary$call_count),
    site_count = sum(count_summary$site_count)
)

count_summary <-
    bind_rows(count_summary, 
              eval_source_type_all_summary, 
              all_eval_function_summary, 
              all_all_summary)

print(count_summary)

count_summary %>%
pmap_dfr(function(eval_source_type, eval_function, call_count, site_count) {
    eval_function <- str_replace(eval_function, "\\.", " ")
    name <- str_c(eval_source_type, " ", eval_function, " ", "call count")
    r(name, call_count)
    name <- str_c(eval_source_type, " ", eval_function, " ", "site count")
    r(name, site_count)
})
```

```{r run-data, eval=TRUE}
run_eval_proportion <-
    all_data %>%
    group_by(file, eval_source_type) %>%
    summarize(eval_count = sum(nb_ev_calls)) %>%
    ungroup() %>%
    pivot_wider(names_from = "eval_source_type", values_from = "eval_count", values_fill = 0) %>%
    mutate(all = core + package)

# add the programs that have no evals
run_eval_proportion <-
  bind_rows(
    run_eval_proportion,
    anti_join(programs, run_eval_proportion, by="file") %>% mutate(core=0, package=0, all=0)
  )

total_file_count <- length(unique(run_eval_proportion$file))
stopifnot(total_file_count == nrow(programs))
# anti_join(run_eval_proportion, programs, by="file")

no_eval_file_count <-
    run_eval_proportion %>%
    filter(core == 0 & package == 0)

core_eval_file_count <-
    run_eval_proportion %>%
    filter(core != 0 & package == 0)

package_eval_file_count <-
    run_eval_proportion %>%
    filter(core == 0 & package != 0)

all_eval_file_count <-
    run_eval_proportion %>%
    filter(core != 0 & package != 0)

overview_table(    
  r("total file count", total_file_count),
  r("no eval file count", no_eval_file_count),
  r("no eval file perc", ratio(no_eval_file_count, total_file_count)),
  r("core eval file count", core_eval_file_count),
  r("core eval file perc", ratio(core_eval_file_count, total_file_count)),
  r("package eval file count", package_eval_file_count),
  r("package eval file perc", ratio(package_eval_file_count, total_file_count)),
  r("all eval file count", all_eval_file_count),
  r("all eval file perc", ratio(all_eval_file_count, total_file_count))
)

run_eval_proportion <-
    run_eval_proportion %>%
    filter(core != 0 & package != 0) %>%
    mutate(core_perc = round(core * 100 / all)) %>%
    group_by(core_perc) %>%
    summarize(file_count = n()) %>%
    ungroup() %>%
    arrange(desc(core_perc)) %>%
    mutate(file_perc = round(100 * file_count / sum(file_count), 1)) %>%
    mutate(cumsum_file_perc = cumsum(file_perc)) %>%
    mutate(cumsum_file_count = cumsum(file_count))

eighty_core_eval_file_count <-
    run_eval_proportion %>%
    filter(core_perc == 80) %>%
    pull(cumsum_file_count)

r("eighty core eval file count", eighty_core_eval_file_count)
r("eighty core eval file perc", ratio(eighty_core_eval_file_count, all_eval_file_count))

run_eval_proportion
```

```{r package-counts, eval=TRUE}
package_calls <-
    packages_data %>%
    filter(eval_source != "base?") %>%
    group_by(eval_source) %>%
    summarize(site_count = length(unique(eval_call_srcref)), 
              call_count = sum(nb_ev_calls)) %>%
    ungroup() %>%
    arrange(desc(call_count)) %>%
    mutate(call_perc = round(100 * call_count / sum(call_count), 1)) %>%
    mutate(cumsum_call_perc = cumsum(call_perc))

top_ten_package_calls <-
    package_calls %>%
    slice(n = 1:10)

datatable(top_ten_package_calls)

top_ten_package_call_count <- sum(top_ten_package_calls$call_count)
top_ten_package_site_count <- sum(top_ten_package_calls$site_count)
total_package_call_count <- sum(package_calls$call_count)

overview_table(
  r("top ten package call count", top_ten_package_call_count),
  r("top ten package call perc", ratio(top_ten_package_call_count, total_package_call_count)),

  r("top ten package site count", top_ten_package_site_count),
  r("top ten package site perc", ratio(top_ten_package_site_count, sum(package_calls$site_count)))
)

top_ten_package_calls %>%
add_column(index = toupper(letters)[1:nrow(top_ten_package_calls)]) %>%
pmap_dfr(function(eval_source, site_count, call_count, call_perc, cumsum_call_perc, index) {
    r(str_c("top ten package name ", index), eval_source)
    r(str_c("top ten package callsite count ", index), site_count)
    r(str_c("top ten package call count ", index), call_count)
    r(str_c("top ten package call perc ", index), ratio(call_count, total_package_call_count))
})
```

```{r callsite-distribution, eval=TRUE}
eval_callsite_table <-
  package_calls %>%
  mutate(
    site_count=cut(
      site_count, 
      breaks=c(0:10, 50, 100, 150, 200, 250, 300, Inf), 
      labels=c(1:10, "11--50", "51--100", "101--150", "151--200", "201--250", "251--300", "> 300"),
      include.lowest = T
    )
  ) %>%
  group_by(site_count) %>%
  summarize(package_count = n())

eval_callsite_table_row <- ceiling(nrow(eval_callsite_table)/3)

eval_callsite_table3 <- map_dfc(seq(1, nrow(eval_callsite_table), eval_callsite_table_row), function(i) {
  x <- tibble(
    eval_callsite_table[i:(i+eval_callsite_table_row), "site_count"],
    eval_callsite_table[i:(i+eval_callsite_table_row), "package_count"]
  )
  colnames(x) <- paste0(colnames(x), i)
  x
})
rownames(eval_callsite_table3) <- NULL

eval_callsite_table3 %>%
  xtable() %>%
  print(
    only.contents=TRUE, 
    include.colnames=FALSE, 
    include.rownames=FALSE,
    hline.after=nrow(eval_callsite_table3),
    file=path(TAGS_DIR, "table-site-package-summary.tex")
  )

datatable(eval_callsite_table3)
```

```{r core-dependent-own-eval}
cod_data <- 
  all_data %>%
  mutate(
    package=basename(dirname(dirname(file))),
    core_own_dependent=case_when(
      eval_source == "core"  ~ "core",
      eval_source == "base"  ~ "core",
      package == eval_source ~ "own",
      TRUE ~ "dependent"
    )
  )
```

```{r}
programs_cod <-
  map_dfr(
    c("core", "own", "dependent"), 
    ~mutate(
      programs,
      core_own_dependent=.,
      nb_ev_calls=0
    )
  )

cod_data_summary <- 
  cod_data %>%
  select(file, core_own_dependent, corpus, nb_ev_calls) %>%
  right_join(programs_cod, by=c("file", "corpus", "core_own_dependent")) %>%
  mutate(nb_ev_calls=ifelse(is.na(nb_ev_calls.x), nb_ev_calls.y, nb_ev_calls.x)) %>%
  select(file, core_own_dependent, corpus, nb_ev_calls)

core_own_dependent <-
  cod_data_summary %>%
  group_by(file, core_own_dependent) %>%
  summarize(corpus=first(corpus), eval_calls = sum(nb_ev_calls)) %>%
  ungroup()

vars <- expand.grid(corpus=c("cran", "kaggle"), cod=c("core", "own", "dependent"), stringsAsFactors = FALSE)
map2_dfr(vars$corpus, vars$cod, function(crps, cod) {
  rs(paste(crps, cod, "eval calls"), filter(core_own_dependent, corpus==crps, core_own_dependent==cod)$eval_calls)
}) %>%
  overview_table()

# core_own_dependent  %>%
#   group_by(corpus, core_own_dependent) %>%
#   pivot_longer(cols=c(total_eval_calls, max_eval_calls, min_eval_calls, avg_eval_calls))
# 
#   xtable::xtable(align="llrrrr") %>%
#   print(
#     tabular.environment = "tabularx", 
#     width = "\\textwidth",
#     floating=F,
# #    file=fname,
#     include.rownames=F
#   )

# core_own_dependent %>%
# pmap_dfr(function(corpus, core_own_dependent, total_eval_calls, max_eval_calls, min_eval_calls, avg_eval_calls) {
#     r(str_c(corpus, " ", core_own_dependent, " total eval calls"), total_eval_calls)
#     r(str_c(corpus, " ", core_own_dependent, " max eval calls"), max_eval_calls)
#     r(str_c(corpus, " ", core_own_dependent, " min eval calls"), min_eval_calls)
#     r(str_c(corpus, " ", core_own_dependent, " avg eval calls"), avg_eval_calls)
# })
# 
# datatable(core_own_dependent)
# 
#     xtable::xtable(align="lXrrrr") %>%
#     print(
#       tabular.environment = "tabularx", 
#       width = "\\textwidth",
#       floating=F,
#       file=fname,
#       include.rownames=F
#     )
```