### 2020-08-28 inst/analysis/patterns.Rmd
Add pattern table

---
title: "Patterns"
output: html_document
params:
  base_dir: ../../../
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
local(for (x in c(
  "dplyr", 
  "DT",
  "fs",
  "fst",
  "ggplot2",
  "ggthemes",
  "lubridate",
  "readr",
  "purrr",
  "stringr",
  "tidyr",
  "tidytext",
  "viridis",
  "xtable"
  )) {
  suppressPackageStartupMessages(library(x, character.only=TRUE))
})

knitr::opts_chunk$set(
  echo = TRUE,
  fig.align = "center",
  fig.retina = 2,
  fig.width = 10
)

# TODO: move into the package
source("inc/paths.R")
source("inc/setup.R")
source("inc/latextags.R")

options(repos="https://cran.r-project.org")
theme_set(theme_minimal())
create_tags(path(TAGS_DIR, "patterns.tex"), prefix="Pattern", default=TRUE)
```

## Data

```{r}
cran_programs <- read_fst(CRAN_PROGRAMS_FILE) %>% as_tibble() %>% filter(success)
kaggle_programs <- read_fst(KAGGLE_PROGRAMS_FILE) %>% as_tibble() %>% filter(success)
programs <- bind_rows(
  select(cran_programs, file, corpus),
  select(kaggle_programs, file, corpus)
)
```

```{r}
core_data <- 
  read_fst(EVALS_SUM_CORE_FILE) %>%
  as_tibble() %>%
  semi_join(programs, by="file")

packages_data <- 
  read_fst(EVALS_SUM_PKGS_FILE) %>%
  as_tibble() %>%
  semi_join(programs, by="file")

# kaggle should be always empty
kaggle_data <- 
  read_fst(EVALS_SUM_KAGGLE_FILE) %>%
  as_tibble() %>%
  semi_join(programs, by="file")

all_data <- 
  bind_rows(core_data, packages_data, kaggle_data) %>%
  filter(eval_function %in% c("eval", "evalq")) %>%
  select(
    file, 
    expr=expr_resolved, 
    expr_type=expr_resolved_type, 
    expr_fun=expr_resolved_function,
    srcref=eval_call_srcref, 
    caller_expression,
    caller_function,
    caller_package,
    expr_parsed_expression, 
    eval_source_type,
    expr_expression_type,
    envir_expression,
    enclos_expression,
    environment_class,
    nb_ev_calls
  )
```

## JS Patterns

```{r}
calls <- function(x) {
  stopifnot("nb_ev_calls" %in% colnames(x))
  sum(x$nb_ev_calls)
}

ratio_calls <- function(x, y) {
  ratio(calls(x), calls(y))
}
```

### library

```{r}
p_lib <- filter(all_data, str_detect(expr, "<-\\s*function\\("))
p_lib_class <- 
  p_lib %>%
  mutate(
    reason=case_when(
      caller_expression == "eval(ei, envir)"    ~ "source", # base/R/source.R:219
      caller_function == "source"               ~ "source",
      caller_function == "sys.source"           ~ "source",
      str_starts(expr, fixed("eval(quote({\n")) ~ "local",
      TRUE                                      ~ "unknown"
    )
  )

count(p_lib_class, reason)
filter(p_lib_class, reason=="unknown") %>% arrange(desc(nb_ev_calls))
```

```{r}
overview_table(
  r("lib",        calls(p_lib_class)),
  r("lib ratio",  ratio_calls(p_lib, all_data)),
  r("lib source", calls(filter(p_lib_class, reason=="source"))),  
  r("lib local",  calls(filter(p_lib_class, reason=="local")))  
)
```

### Read

resolving variable -- `evalq(x)` or `eval(quote(x))`

```{r}
p_read <-
  all_data %>%
  filter(expr_type=="SYMSXP" | expr_fun %in% c("[", "[[", "$", "slot", "@", "::", ":::"))

p_read %>% 
  count(caller_function, sort=T) %>% print(n=Inf)

p_read_slot_access <- filter(p_read, expr_fun %in% c("[", "[[", "$", "slot", "@", "::", ":::"))

overview_table(
  r("read", calls(p_read)),
  r("read ratio", ratio_calls(p_read, all_data)),
  r("read slot access ratio", ratio_calls(p_read_slot_access, p_read)),
  r("read datatable ratio", ratio_calls(filter(p_read, caller_package=="data.table"), p_read_slot_access))
)
```

### Assign

```{r}
p_assign <-
  all_data %>%
  filter(expr_fun == "=" | str_ends(expr_fun, "<-") | expr_fun %in% c("assign", "delayedAssign"))

p_assign_arrow <- filter(p_assign, expr_fun=="<-")
p_assign_superarrow <- filter(p_assign, expr_fun=="<<-")

overview_table(
  r("assign", calls(p_assign)),
  r("assign ratio", ratio_calls(p_assign, all_data)),
  r("assign arrow ratio", ratio_calls(p_assign_arrow, p_assign)),
  r("assign arrow magrittr ratio", ratio_calls(filter(p_assign_arrow, caller_package=="magrittr"), p_assign_arrow)),
  r("assign superarrow ratio", ratio_calls(p_assign_superarrow, p_assign)),
  r("assign superarrow plyr ratio", ratio_calls(filter(p_assign_superarrow, caller_package=="plyr"), p_assign_superarrow))
)

count(p_assign, expr_fun) %>%
  rowwise() %>%
  do({
    e_f <- .$expr_fun
    e_n <- .$n
    tibble(e_f, e_n, calls=calls(filter(p_assign, expr_fun==e_f)))
  })

count(p_assign, caller_package) %>%
  rowwise() %>%
  do({
    e_p <- .$caller_package
    e_n <- .$n
    tibble(e_p, e_n, calls=calls(filter(p_assign, caller_package==e_p)))
  })
```

### typeof

```{r}
p_typeof <- 
  all_data %>%
  filter(
    expr_fun=="typeof" |
    expr_fun=="class" |
    str_starts(expr_fun, "is\\.")
  )

count(p_typeof, expr_fun) %>%
  rowwise() %>%
  do({
    e_f <- .$expr_fun
    e_n <- .$n
    tibble(e_f, e_n, calls=calls(filter(p_typeof, expr_fun==e_f)))
  })

overview_table(
  r("typeof", calls(p_typeof)),
  r("typeof ratio", ratio_calls(p_assign, all_data)),
  r("assign arrow ratio", ratio_calls(filter(p_assign, expr_fun=="<-"), p_assign)),
  r("assign superarrow ratio", ratio_calls(p_assign_superarrow, p_assign)),
  r("typeof assertthat ratio", ratio_calls(filter(p_assign_superarrow, caller_package=="plyr"), p_assign_superarrow))
)
```

### try

```{r}
p_try <- 
  all_data %>%
  filter(
    expr_fun %in% c("try", "tryCatch")
  )

overview_table(
  r("try", calls(p_try)),
  r("try ratio", ratio_calls(p_try, all_data))
)
```

mostly from methods

### empty

```{r}
p_empty <- all_data %>%
  filter(expr=="" | str_detect(expr, "^[\"']\\s*[\"']$"))

overview_table(
  r("empty", calls(p_empty)),
  r("empty ratio", ratio_calls(p_empty, all_data))
)
```

### call

```{r}
p_call <- all_data %>%
  filter(
    expr_type %in% c("LANGSXP", "EXPRSXP")
  ) %>%
  anti_join(
    bind_rows(
      p_lib, p_read, p_assign, p_typeof, p_try, p_empty
    )
  )

filter(all_data, expr_type == "EXPRSXP", !str_starts(expr, "^.*\\(")) %>% count(expr)

overview_table(
  r("call", calls(p_call)),
  r("call ratio", ratio_calls(p_call, all_data))
)
```

### other

```{r}
p_other <-
  anti_join(
    all_data,
    bind_rows(
      p_lib, p_read, p_assign, p_typeof, p_try, p_call, p_empty
    )
  )

overview_table(
  r("other", calls(p_other)),
  r("other ratio", ratio_calls(p_other, all_data))
)
```

```{r}
merge_patterns <- function(...) {
  vars <- substitute(list(...))
  vars <- as.list(vars[-1])

  map_dfr(
    vars,
    function(x) {
      df <- get(as.character(x))
      mutate(df, pattern=as.character(x))
    }
  )
}

merge_patterns_sum <- function(...) {
  vars <- substitute(list(...))
  vars <- as.list(vars[-1])

  all <- all_data %>% group_by(eval_source_type) %>% summarize(all_evals=sum(nb_ev_calls))

  map_dfr(
    vars,
    function(x) {
      name <- as.character(x)
      df <- get(name)
      if (nrow(df) == 0) {
        tibble(eval_source_type=all$eval_source_type, pattern=name, nb_ev_calls=0, ratio=0)
      } else {
        df %>%
          mutate(pattern=name) %>%
          select(eval_source_type, nb_ev_calls, pattern) %>%
          group_by(eval_source_type, pattern) %>%
          summarize_all(sum) %>%
          left_join(all, by="eval_source_type") %>%
          mutate(ratio=nb_ev_calls/all_evals) %>%
          select(-all_evals)
      }
    }
  )
}
```

```{r}
patterns <- merge_patterns(p_lib, p_read, p_assign, p_typeof, p_try, p_call, p_other)

patterns_sum <- merge_patterns_sum(p_lib, p_read, p_assign, p_typeof, p_try, p_call, p_other)
```

```{r}
patterns_sum %>%
  group_by(eval_source_type) %>%
  summarise(ratio=sum(ratio))
```

```{r}
patterns_sum %>%
  ggplot(aes(x=reorder(pattern, -ratio), y=ratio)) +
  geom_bar(stat="identity") +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~eval_source_type)
```

```{r}
patterns_table <- 
  patterns_sum %>%
  mutate(ratio=map_chr(ratio, ~fmt(percent(.))), nb_ev_calls=map_chr(nb_ev_calls, ~fmt(oom(.)))) %>%
  pivot_wider(names_from=eval_source_type, values_from=c(nb_ev_calls, ratio), values_fill=list(nb_ev_calls="0", ratio="0%")) %>%
  left_join(
    tibble(
      name   =c("Library", "Read", "Assign",   "Typeof",   "Try",   "Call",   "Empty",   "Other"),
      pattern=c("p_lib", "p_read", "p_assign", "p_typeof", "p_try", "p_call", "p_empty", "p_other")
    ),
    by="pattern"
  ) %>%
  select(name, nb_ev_calls_core, ratio_core, nb_ev_calls_package, ratio_package)

patterns_table %>%
  arrange(desc(ratio_core)) %>%
  xtable() %>%
  print(
    only.contents=TRUE, 
    include.colnames=FALSE, 
    include.rownames=FALSE,
    hline.after=nrow(patterns_table),
    file=path(TAGS_DIR, "table-js-patterns-in-r.tex")
  )
```

```
p_assign_ast <-
  p_assign %>%
  mutate(expr_ast=map(expr, ~tryCatch(str2lang(.), error=function(e) NA)))

p_assign_ast_len <-
  p_assign_ast %>%
  mutate(expr_ast_length=map_int(expr_ast, length)) %>%
  filter(expr_ast_length==3) %>%
  mutate(fun=map_chr(expr_ast, function(x) {
    if (length(x[[1]]) == 1) as.character(x[[1]]) else NA
  })) %>%
  filter(!is.na(fun))

p_assign_arrow <- 
  filter(p_assign_ast_len, fun %in% c("<-", "<<-"))

p_assign_arrow %>% count(caller_function, sort = TRUE)

overview_table(
  r("assign", p_assign_arrow),
  r("assign ratio", ratio(p_assign_arrow, all_data)),
  r("assign super", ratio(filter(p_assign_ast_len, fun == "<<-"), p_assign_ast_len)),
  r("assign maggit", ratio(filter(p_assign_ast_len, caller_function == "%<>%"), p_assign_ast_len))
)
```

## Patterns

### Value

```{r}

```

```
filter(corpus, evals>0)$package %>% sample(20)
```

```{r}
packages <- c(
  "timeSeries",
  "irlba",
  "assertthat",
  "rgenoud",
  "taxize",
  "prodlim",
  "vars" ,
  "betareg"   ,
  "bookdown"  ,
  "R2HTML"   ,
  "withr",
  "processx" ,
  "modeltools"  ,
  "rprojroot"  ,
  "ggplot2",
  "binom",
  "rmarkdown" ,
  "R.methodsS3" ,
  "FactoMineR" ,
  "PerformanceAnalytics"
)
```

```{r}
evals_static <- read_fst(PACKAGE_EVALS_FILE)
```

### parse

```{r}
all_data %>%
  filter(!is.na(parse_args_text))
```

### timeSeries

```{r}
filter(evals_static, package=="timeSeries")
```

```
as.timeSeries.character - dynamic code loading
    # Load Demo File - Returns a data frame:
    x <- eval(parse(text = eval(x)))

setMethod("rank", "timeSeries",
  function(x,  na.last = TRUE,
      ties.method = eval(formals(rank)$ties.method)) -- incorrect usage

```

```{r}
paste0("scp -r prl3:/var/lib/R/project-evalR/CRAN/extracted/",packages," . &&") %>% cat(sep="\n")
```