### 2020-11-03 inst/analysis/environments.Rmd
Classification of environments

---
title: "Empirical Evaluation of eval in R"
author: "Pierre Donat-Bouillud"
output:
  rmdformats::readthedown:
        code_folding: hide
        lightbox: true
        gallery: false
        df_print: paged
        toc_depth: 3
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
params:
  base_dir: /var/lib/R/project-evalR
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.align = "center",
                      fig.retina = 2,
                      fig.width = 10,
                      cache.lazy = FALSE)

now <- Sys.time()

library(tidyverse)
library(fst)

source("insights.R", local = knitr::knit_global())
source("preprocess.R", local = knitr::knit_global())
```

# Loading datafiles

```{r loading_fst}
eval_calls <- read_fst("/var/lib/R/project-evalR/run/package-evals-traced-test/calls.fst") %>%
  as_tibble() %>%
  add_types()
```

# Description of the data files

```{r packages}

```

# Classification of environments

We distinguish between 3 main categories of `eval` calls with respect to their `envir` argument:

- `envir` is a list or a data frame
- `envir` comes from an argument of the caller function
- `envir` is absolutely computed (with `parent.frame()` or equivalent)

##  `envir` as a list or data frame

```{r envir_list}
eval_calls_env_list <- eval_calls %>% filter(envir_type == "VECSXP")
```

## `envir` as an argument from the caller function

```{r envir_arg_caller}
eval_calls_arg_caller <- eval_calls %>% filter(!is.na(envir_from_arg))
```

## `envir` absolutely computed

## `envir` absolutely computed

```{r}
duration <- difftime(Sys.time(), now)
```

Notebook execution was `r duration` `r units(duration)`.