### 2020-12-15 analysis/corpus-stage3.Rmd
changes!

---
title: "Corpus stage 3"
output: html_document
params:
  base_dir: /var/lib/R/project-evalR
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
for (x in c(
  "dplyr", 
  "DT",
  "fs",
  "fst",
  "lubridate",
  "readr",
  "purrr",
  "stringr",
  "tidyr"
  )) {
  suppressPackageStartupMessages(library(x, character.only=TRUE))
}

knitr::opts_chunk$set(echo = TRUE)

# TODO: move into the package
source("inc/paths.R")
source("inc/setup.R")
```

## Load data

```{r}
calls       <- read_fst(CALLS_FILE) %>% mutate(package=basename(dirname(dirname(file))))
corpus      <- read_fst(CORPUS_DETAILS_FILE)
calls_final <- semi_join(calls, corpus, by="package")
```

```{r export data}
write_fst(calls_final, "../../run/package-evals-traced/calls-final.fst")
```