### 2020-12-15 analysis/side-effects.Rmd

---
title: "Side effects"
output: html_document
params:
  base_dir: /home/aviral/projects/revalstudy2
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
for (x in c(
  "dplyr", 
  "DT",
  "fs",
  "fst",
  "ggplot2",
  "ggthemes",
  "lubridate",
  "readr",
  "purrr",
  "stringr",
  "tidyr"
  )) {
  suppressPackageStartupMessages(library(x, character.only=TRUE))
}

knitr::opts_chunk$set(
  echo = TRUE,
  fig.align = "center",
  fig.retina = 2,
  fig.width = 10
)

# TODO: move into the package
source("inc/paths.R")
source("inc/setup.R")
source("inc/latextags.R")

options(repos="https://cran.r-project.org")
theme_set(theme_minimal())
create_tags(path(TAGS_DIR, "side-effects.tex"), prefix="", default=TRUE)
```

```{r}
core_data <- read_fst(EVALS_SUM_CORE_FILE)
package_data <- read_fst(EVALS_SUM_PKGS_FILE)
kaggle_data <- read_fst(EVALS_SUM_KAGGLE_FILE)
program_data <- read_fst(PROGRAM_FILE)

all_data <- rbind(core_data, package_data, kaggle_data)
```

## Side-Effects and Eval

Each entry in `call.fst` contains `direct_writes` and `indirect_writes` (both
`0` for `eval.parent` and `local`). `direct_writes` are the side-effects
directly caused by the `eval` in question and `indirect_writes` are the
side-effects caused by all children evals of the eval. All application
assignments can be found in `direct_writes` field of `program.fst` and all
`eval_assignments` can be found in `indirect_writes` field of `program.fst`. All
non-local assignments can be found by simply summing up the `direct_writes`
field of all eval calls.

```{r preprocess-write-data}
eval_writes_data <-
    all_data %>%
    filter(eval_function == "eval" | eval_function == "evalq") %>%
    select(eval_source_type, eval_source, eval_call_srcref, direct_writes, indirect_writes, nb_ev_calls) %>%
    mutate(all_writes = direct_writes + indirect_writes, side_effecting = as.logical(all_writes))

program_writes_data <-
    program_data %>%
    select(direct_writes, indirect_writes)
```

```{r compute-side-effecting-write-counts}
all_writes <- sum(program_writes_data$direct_writes)
eval_writes <- sum(program_writes_data$indirect_writes)

r("All Writes", all_writes)
r("Eval Writes", eval_writes)

eval_side_effecting_writes <- sum(eval_writes_data$direct_writes * eval_writes_data$nb_ev_calls)

r("Eval Side Effecting Writes", eval_side_effecting_writes)

r("Eval Side Effecting Writes Eval Perc", ratio(eval_side_effecting_writes, eval_writes))
r("Eval Side Effecting Writes All Perc", ratio(eval_side_effecting_writes, all_writes))
```

```{r compute-side-effecting-call-counts}
eval_calls_by_source_writes <-
    eval_writes_data %>%
    count(eval_source_type, side_effecting, wt = nb_ev_calls, name = "call_count") %>%
    group_by(eval_source_type) %>%
    mutate(total_call_count = sum(call_count)) %>%
    ungroup() %>%
    mutate(call_perc = round(100 * call_count/total_call_count, 1)) %>%
    filter(side_effecting)

datatable(eval_calls_by_source_writes)

eval_calls_by_source_writes %>%
pmap(function(eval_source_type, side_effecting, call_count, total_call_count, call_perc) {
    r(str_c("Side Effecting ", eval_source_type, " call perc"), str_c(call_perc, "%"))
})

eval_calls_by_package_writes <-
    eval_writes_data %>%
    filter(eval_source_type != "core") %>%
    count(eval_source, side_effecting, wt = nb_ev_calls, name = "call_count") %>%
    arrange(desc(call_count))

datatable(eval_calls_by_package_writes)

eval_calls_by_source_write_count <-
    eval_writes_data %>%
    filter(all_writes != 0) %>%
    mutate(all_writes = if_else(all_writes <= 5, as.character(all_writes), "> 5")) %>%
    count(eval_source_type, all_writes, wt = nb_ev_calls, name = "call_count") %>%
    group_by(eval_source_type) %>%
    mutate(total_call_count = sum(call_count)) %>%
    ungroup() %>%
    mutate(call_perc = round(100 * call_count/total_call_count, 1))

datatable(eval_calls_by_source_write_count)

plot <-
    eval_calls_by_source_write_count %>%
    ggplot(aes(factor(all_writes, level = c(1:5, "> 5")), call_perc)) +
    geom_col(aes(fill = eval_source_type), position = "dodge") +
    labs(title="", x ="Side-Effects", y = "Side-Effecting Eval Calls %", fill = "") +
    theme(text = element_text(size=15), axis.text.x = element_text(size=15))
plot
```

The call site computation is incorrect.

```{r compute-side-effecting-site-counts, eval = FALSE}
eval_sites_by_source_writes <-
    eval_writes_data %>%
    group_by(eval_source_type, eval_call_srcref) %>%
    summarize(side_effecting = any(side_effecting)) %>%
    ungroup() %>%
    count(eval_source_type, side_effecting, name = "site_count") %>%
    group_by(eval_source_type) %>%
    mutate(total_site_count = sum(site_count)) %>%
    ungroup() %>%
    mutate(site_perc = round(100 * site_count / total_site_count, 1)) %>%
    filter(side_effecting)

datatable(eval_sites_by_source_writes)

eval_sites_by_source_writes %>%
pmap(function(eval_source_type, side_effecting, site_count, total_site_count, site_perc) {
    r(str_c("SideEffecting ", eval_source_type, " site perc"), str_c(site_perc, "%"))
})

eval_sites_by_package_writes <-
    eval_writes_data %>%
    filter(eval_source_type != "core") %>%
    group_by(eval_source, eval_call_srcref) %>%
    summarize(side_effecting = any(side_effecting)) %>%
    ungroup() %>%
    count(eval_source, side_effecting, name = "site_count") %>%
    group_by(eval_source, side_effecting) %>%
    mutate(site_perc = round(100 * site_count / sum(site_count), 1)) %>%
    ungroup() %>%
    arrange(side_effecting, desc(site_perc))

datatable(eval_sites_by_package_writes)
```