### 2020-12-21 analysis/usage.Rmd
minor

---
title: "Usage metrics"
author: "Pierre Donat-Bouillud"
output:
    html_document:
        gallery: false
        toc: true
        toc_depth: 3
        toc_float: true
        df_print: paged
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
params:
  base_dir: Data/
  calls_path: package/summarized.fst
  corpus_path: corpus.fst
  static_path: package-evals-static.csv
  force_rebuild: FALSE
---

```{r setup, include=FALSE}
library(tidyverse)
library(fst)
library(fs)
library(DT)
library(scales)

knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) { # record the current time before each chunk
      now <<- Sys.time()
    } else { # calculate the time difference after a chunk
      res <- difftime(Sys.time(), now)
      # return a character string to show the time
      paste("Time for this code chunk to run:", round(res, 2), units(res))
    }
  }
}))
now <- Sys.time()
knitr::opts_chunk$set(
  echo = TRUE,
  fig.retina = 2,
  fig.width = 10,
  cache.lazy = FALSE,
  time_it = TRUE
)

source("inc/paths.R")
source("inc/latextags.R")
source("inc/setup.R")
theme_set(theme_minimal())
create_tags(path(TAGS_DIR, "usage_metrics.tex"), prefix = "", default = TRUE)

calls_path <- paste0(params$base_dir, params$calls_path)
corpus_path <- paste0(params$base_dir, params$corpus_path)
static_path <- paste0(params$base_dir, params$static_path)
```

# Read

Read in the summarized calls data (`E_raw`) and the file that describes
which package belong to our corpus (`C_raw`). For conciseness, `E` is
the input data set with only the columns that we use.

```{r read, message=TRUE, include=F}
read_fst(corpus_path) %>% as_tibble() -> C_raw
read_csv(static_path) %>% semi_join(C_raw, by = "package") -> P

rebuild <- TRUE
# We only read the file if we don't have our summarized file on disk.
# To force reading, set the force_rebuild parameter to true.
if (file.exists("Data/.lock.fst")) {
  saved <- read.fst("Data/.lock.fst")[[1,1]]
  if (saved == file.size(calls_path)) {
    read.fst("Data/E.fst") %>% as_tibble() -> E
    rebuild <- FALSE
  }
}
if (params$force_rebuild | rebuild) {
  read_fst(calls_path) %>% as_tibble() -> E_raw

  if (!"direct_interpreter_eval" %in% names(E_raw)) { # We are dealing with
    E_raw %>% mutate(direct_interpreter_eval=0) -> E_raw # partial data.
  } #                                                  We create a fake column.

  E_raw %>% select(
    run_package = package, #      The package which was run to trigger the eval
    ev_variant = eval_function, # Which eval variant was called
    duplicates = nb_ev_calls, #   Number of identical eval calls (weight)
    src_ref = eval_call_srcref, # Source ref for the eval call site
    file, #                       Script file name (name of this run)
    ev_package = eval_source, #   Package of the call site
    arg_size = expr_resolved_length, #   Size in characters of argument 
    ast_nodes = expr_resolved_nodes, #   Size in AST Nodes
    type = expr_resolved_type_tag,   #   Argument type
    opcodes = direct_interpreter_eval #  Count of the AST nodes traversed by the
  ) %>% #                                interpreter during eval (non-recursive).
   mutate(across(where(is_character),as_factor)) -> E 
  write.fst(E, "Data/E.fst")
  sz <- file.size(calls_path) %>% as_tibble()
  write.fst(sz,"Data/.lock.fst")
}

C_raw -> C
corpus_size <- nrow(C)

nb_runs <- E %>%
  pull(file) %>%
  n_distinct()

overview_table(
  r("Corpus", corpus_size),
  r("NbRuns", nb_runs)
)
```

The two data sets are `E` and `C` for conciseness.  `C` is only used here to 
compute the corpus size. `E` is used throughout. Recall that `E_raw` has the 
source data with all columns. `P` corresponds to static information about the
call sites.

The size of our corpus is `r corpus_size` packages.

#  Variants

The data set has `r nrow(E)` distinct calls to the `eval` and its variants. 
There are `r length(unique(E$ev_package))` corpus packages with `eval` calls
(note that we may not trigger all `eval`s in a package due to spotty coverage).

```{r variants, include=FALSE}
nb_calls <- E %>%
  count(wt = duplicates) %>%
  .$n
eval_calls <- E %>%
  filter(ev_variant == "eval") %>%
  count(wt = duplicates) %>%
  .$n
evalq_calls <- E %>%
  filter(ev_variant == "evalq") %>%
  count(wt = duplicates) %>%
  .$n
eval.parent_calls <- E %>%
  filter(ev_variant == "eval.parent") %>%
  count(wt = duplicates) %>%
  .$n
local_calls <- E %>%
  filter(ev_variant == "local") %>%
  count(wt = duplicates) %>%
  .$n

overview_table(
  r("Allcalls", nb_calls),
  r("Evals", eval_calls),
  r("Evalqs", evalq_calls),
  r("Eparents", eval.parent_calls),
  r("Locals", local_calls),
  r("Triggeredpkgs", length(unique(E$ev_package)))
) -> TBL
```

We recorded `r nb_calls` to eval, out of these, there were `r evalq_calls`
(`r fmt(ratio(evalq_calls,nb_calls))`) calls to `evalq`, `r eval.parent_calls` 
(`r fmt(ratio(eval.parent_calls,nb_calls))`) calls to `eval.parent`,  and
`r local_calls` (`r fmt(ratio(local_calls,nb_calls))`) calls to `local`.

`r TBL`

# Exercised vs. Static

Here we compute the number of static call sites to compare it with
the number of call sites with triggered dynamically. The former is larger
than the latter because code coverage is not perfect.

```{r exercised-vs-static, include=FALSE}
stat_evl <- P %>% filter(call_fun_name == "eval")        %>% nrow()
stat_evq <- P %>% filter(call_fun_name == "evalq")       %>% nrow()
stat_evp <- P %>% filter(call_fun_name == "eval.parent") %>% nrow()
stat_loc <- P %>% filter(call_fun_name == "local")       %>% nrow()

sites_per_variant <- E %>% # src_ref indicates where the call originates from.
  group_by(ev_variant) %>%
  summarize(n = n_distinct(src_ref)) # This the number of distinct sites

tmp <- sites_per_variant %>% filter(ev_variant == "eval")
dyn_evl <- if (nrow(tmp) == 0) { 0 } else { tmp$n }
tmp <- sites_per_variant %>% filter(ev_variant == "evalq")
dyn_evq <- if (nrow(tmp) == 0) { 0 } else { tmp$n }
tmp <- sites_per_variant %>% filter(ev_variant == "eval.parent")
dyn_evp <- if (nrow(tmp) == 0) { 0 } else { tmp$n }
tmp <- sites_per_variant %>% filter(ev_variant == "local")
dyn_loc <- if (nrow(tmp) == 0) { 0 } else { tmp$n }

overview_table(
  r("Staticeval", stat_evl),
  r("Staticevalq", stat_evq),
  r("Staticevalparent", stat_evp),
  r("Staticlocal", stat_loc),
  r("Triggeredeval", dyn_evl),
  r("Triggeredevalq", dyn_evq),
  r("Triggeredevalparent", dyn_evp),
  r("Triggeredlocal", dyn_loc),
  r("Triggeredevalpct", ratio(dyn_evl, stat_evl)),
  r("Triggeredevalqpct", ratio(dyn_evq, stat_evq)),
  r("Triggeredevalparentpct", ratio(dyn_evp, stat_evp)),
  r("Triggeredlocalpct", ratio(dyn_loc, stat_loc))
) -> TBL
```

`r TBL`

# Ev-dist

We compute the frequency of evals by package. This shows that some packages
hardly use eval while other use it intensly. The numbers are "for all runs",
this means that a popular package may be triggered by more runs and thus
have higher frequencies. We will later control for this by looking at individual 
call sites.

```{r ev-dist, include=FALSE}

# The calls per package
pack_calls <- E %>% count(ev_package, wt = duplicates, sort = TRUE)
# Create bins at every power of ten
bins <- cut(pack_calls$n, breaks = c(0, 10^seq(1, 8)), include.lowest = TRUE)
sum <- summary(bins)
# Packages with more than 1K calls
Kcalls <- pack_calls %>%
  filter(n > 1000) %>%
  nrow()

# The number of runs per package
runs <- E %>%
  group_by(ev_package) %>%
  summarise(nb_runs = n_distinct(file))

# Number of eval calls per package
ecpp <- pack_calls %>% left_join(runs, by = "ev_package")
stopifnot(nrow(filter(ecpp, n == 0)) == 0)
# The package with the most calls
max_calls <- ecpp %>% filter(n == max(n))

overview_table(
  r("bina", sum[[1]]),
  r("binb", sum[[2]]),
  r("binc", sum[[3]]),
  r("bind", sum[[4]]),
  r("bine", sum[[5]]),
  r("binf", sum[[6]]),
  r("bing", sum[[7]]),
  r("binh", sum[[8]]),
  r("Fewcalls", sum[[1]] + sum[[2]]),
  r("Manycalls", Kcalls),
  r("Maxcalls", max_calls$n),
  r("Maxcallspack", max_calls$ev_package)
) -> TBL
```

The distribution of eval calls, each bin represents a power of 10, the number 
in the bin is the count of packages.

`r TBL`

# Triggered sites

```{r triggered-sites, include=F}
stopifnot(sum(is.na(E$src_ref)) == 0) # No NAs in scr_ref!
# Number of eval call sites
sites <- E %>% # src_ref indicates where the call originates from.
  pull(src_ref) %>%
  n_distinct() # This the number of distinct sites
# Number of call sites per package
sites_per_package <- E %>%
  count(ev_package, src_ref) %>%
  count(ev_package) # All packages have at least one site!

static_sites <- nrow(P)

static_call_sites_per_package <- P %>%
  count(package, sort = TRUE)

static_at_least_one_call_site <- static_call_sites_per_package %>%
  filter(n >= 1)

overview_table(
  r("Sites", sites),
  r("Staticsites", static_sites),
  r("Sitespercent", ratio(sites, static_sites)),
  r("Staticatleastonecallsite", nrow(static_at_least_one_call_site))
)
```

`r TBL`

# Numbers of calls to eval

This is the number of eval calls which have a call site in each package,
divided by the number of runs for the package.

We first compute an approximation of the number of runs per package.
```{r runs-per-package, include=FALSE}
runs_package <- E %>%
  group_by(ev_package) %>%
  summarise(nb_runs = n_distinct(file)) %>%
  arrange(desc(nb_runs))
```

What are the packages with the most runs?

```{r show_run-per-package}
runs_package %>% datatable()
```

```{r eval-calls-violin-plot}
sites_per_package %>%
  ggplot(aes(x = "", y = n, fill = ""), na.rm = TRUE) +
  geom_violin(trim = FALSE, width = 1.7) +
  labs(y = "Eval calls per package") +
  theme(legend.position = "none")

ggsave(path(PLOT_DIR, "eval_calls_per_packages.pdf"))
```

# Amount of code

How much code is loaded by eval, we look at argument sizes.

```{r amout-of-code, include=F}

weighted.median <- function(x, w, na.rm = FALSE) {
  if (na.rm) {
    na.omit(x)
  }
  w <- w[order(x)]
  x <- x[order(x)]
  prob <- cumsum(w) / sum(w)
  ps <- which(abs(prob - .5) == min(abs(prob - .5)))
  return(x[ps])
}

med_size <- weighted.median(E$ast_nodes, w = E$duplicates, na.rm = TRUE)
max_size <- max(E$ast_nodes, na.rm = TRUE)
average_size <- weighted.mean(E$ast_nodes, na.rm = TRUE, w = E$duplicates)

sizes_prop <- E %>%
  count(ast_nodes, wt = duplicates) %>%
  mutate(proportion = n / sum(n)) %>%
  arrange(ast_nodes) %>%
  mutate(cumulative = cumsum(proportion))

many <- sizes_prop %>%
  filter(cumulative >= 0.95) %>%
  .[1, ]

overview_table(
  r("Medianszeval", med_size),
  r("Maxszeval", max_size),
  r("Avgszeval", average_size),
  r("Szninefive", many$ast_nodes)
) -> TBL
```

`r TBL`

```{r echo=FALSE}
lbl_colour <- rep("black", 11)
lbl_colour[[4]] <- "transparent"
E %>%
  ggplot(aes(x = ast_nodes, weight = duplicates)) +
  stat_density(aes(y = ..count..), color = "black", fill = "blue", alpha = 0.3) +
  scale_x_continuous(breaks = c(0, 1, 2, 3, 4, 5, 10, 30, 100, 300, 1000), trans = "log1p", expand = c(0, 0)) +
  labs(x = "Size (nodes)") +
  theme(text = element_text(size=17),
        axis.text.x = element_text(color = lbl_colour))

E %>% 
  filter(ast_nodes < 25) %>%
  ggplot(aes(x = ast_nodes, weight = duplicates)) +
  geom_histogram(binwidth = 1, color = "black", fill = "blue", alpha = 0.3) +
  scale_y_continuous(labels = comma) +
  labs(x = "nodes") +
  theme(text = element_text(size=17),  panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(), aspect.ratio = 4/3)

ggsave(path(PLOT_DIR, "size_loaded_distribution.pdf"))
```

# Amount of computation

How much work is involved in each eval as measured by the number of opcodes exectued by the interpreter.

```{r amount_computations, include=F}
tot_events <- E %>%
  summarise(n = sum(opcodes * duplicates)) %>%
  .$n
max_events <- E %>%
  summarize(n = max(opcodes)) %>%
  .$n

less_50_events <- E %>%
  filter(opcodes <= 50) %>%
  count(wt = duplicates) %>%
  .$n

overview_table(
  r("Events", tot_events),
  r("Maxevents", max_events),
  r("Smalleventspct", ratio(less_50_events, nb_calls))
) -> TBL
```

`r TBL`

```{r echo=FALSE}
E %>%
  filter(opcodes > 50) %>%
  ggplot(aes(x = "", y = opcodes), na.rm = TRUE) +
  geom_violin(trim = FALSE, width = 1, fill = "blue", alpha = 0.3) +
  scale_y_log10(labels = scales::comma) +
  labs(y = "", x = "") +
  theme(legend.position = "none", text = element_text(size=25))

ggsave(path(PLOT_DIR, "events_per_pack_large.pdf"))
```

```{r echo=FALSE}
E %>%
  filter(opcodes > 0) %>% # == 0 if from eval.parent, because we do not count eval loops for it
  filter(opcodes <= 50) %>%
  ggplot(aes(x = "", y = opcodes), na.rm = TRUE) +
  geom_violin(trim = FALSE, width = 1.7, fill = "blue", alpha = 0.3) +
  labs(y = "Instructions", x = "") +
  theme(legend.position = "none", text = element_text(size=25))

ggsave(path(PLOT_DIR, "events_per_pack_small.pdf"))
```

# Loops

Here, we try to have an estimation of how often a call site is exercised per 
run. If it is often, maybe it is a loop!

```{r loops, echo=FALSE}
# Which call sites where exercised and how many times, per run.
exercised <- E %>%
  group_by(file) %>% # group by runs
  count(src_ref, wt = duplicates) # weigh by duplicates

exercised %>% ggplot(aes(n)) +
  geom_histogram() +
  scale_x_log10(labels = scales::comma) +
  labs(x = "Calls per site per run")

# For each site, how many runs and within each runs how often do we trigger.
triggered <- E %>%
  group_by(src_ref) %>%
  summarize(
    n_runs = n_distinct(file),
    n = sum(duplicates),
    average = n / n_runs
  ) %>%
  arrange(desc(average))

# Plot the low trigger counts...
triggered %>%
  filter(average < 50) %>%
  ggplot(aes(x = average)) +
  geom_histogram(binwidth = 1, fill = "blue", alpha = 0.3, colour = "black") +
  theme(text = element_text(size=16), 
        axis.text = element_text(size=20), 
        aspect.ratio = 4/3)

ggsave(path(PLOT_DIR, "calls_per_run_per_call_site.pdf"))
```

```{r, include = FALSE}
bins <- cut(triggered$average, breaks = c(0, 50, 100, 250, 500, 1000, 1500, 2000, 3000), include.lowest = TRUE)
sum <- summary(bins)

overview_table(
  r("Runbina", sum[[1]]),
  r("Runbinb", sum[[2]]),
  r("Runbinc", sum[[3]]),
  r("Runbind", sum[[4]]),
  r("Runbine", sum[[5]]),
  r("Runbinf", sum[[6]]),
  r("Runbing", sum[[7]]),
  r("Runbinh", sum[[8]])
) -> TBL
```

`r TBL`

# Diversity of arguments

How diverse are the expressions passed to `eval`? We consider the diversity in types in the resolved expression.

```{r type-diversity, echo=FALSE}
E %>%
  ggplot(aes(x = fct_rev(fct_infreq(type)), weight=duplicates)) +
  geom_bar() +
  coord_flip() +
  labs(x = "type")

types <- E %>% count(type, wt=duplicates)

symbs <- types %>% filter(type == "symbol") %>% .$n
languages <- types %>% filter(type == "language") %>% .$n
expressions <- types %>% filter(type == "expression") %>% .$n

ggplot2_symbs <- E %>% filter(ev_package == "ggplot2") %>% 
  count(type, wt=duplicates) %>% filter(type == "symbol") %>% .$n

overview_table(
  r("Symbolpercent", ratio(symbs, nb_calls)),
  r("Languagepercent", ratio(languages, nb_calls)),
  r("Expressionpercent", ratio(expressions, nb_calls)),
  r("Codepercent", ratio(symbs + languages + expressions, nb_calls)),
  r("Ggplotsymbolpercent", ratio(ggplot2_symbs, symbs))
) -> TBL
```

`r TBL`

```{r include=FALSE}
duration <- difftime(Sys.time(), now)
```

Notebook execution was `r duration` `r units(duration)`.