### 2021-01-19 analysis/expressions.Rmd
Canonical expression

---
title: "Eval expressions"
output:
    html_document:
        gallery: false
        toc: true
        toc_depth: 3
        toc_float: true
        df_print: paged
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
params:
  base_dir: Data/package/
  calls_path: summarized.fst
  corpus_path: ../corpus.fst
  static_path: evals-static.csv
  force_rebuild: TRUE
---

```{r setup, include=FALSE}
library(tidyverse)
library(fst)
library(fs)
library(DT)
library(scales)

knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) { # record the current time before each chunk
      now <<- Sys.time()
    } else { # calculate the time difference after a chunk
      res <- difftime(Sys.time(), now)
      # return a character string to show the time
      paste("Time for this code chunk to run:", round(res, 2), units(res))
    }
  }
}))
now <- Sys.time()
knitr::opts_chunk$set(
  echo = TRUE,
  fig.retina = 2,
  fig.width = 10,
  cache.lazy = FALSE,
  time_it = TRUE
)

source("inc/paths.R")
source("inc/latextags.R")
source("inc/functions.R")
theme_set(theme_minimal())
create_tags(path(TAGS_DIR, "usage_metrics.tex"), prefix = "", default = TRUE)

calls_path <- paste0(params$base_dir, params$calls_path)
corpus_path <- paste0(params$base_dir, params$corpus_path)
static_path <- paste0(params$base_dir, params$static_path)
```

# Read

Read in the summarized calls data (`E_raw`) and the file that describes
which package belong to our corpus (`C_raw`). For conciseness, `E` is
the input data set with only the columns that we use.

```{r read, message=TRUE, include=F}
read_fst(corpus_path) %>% as_tibble() -> C_raw
read_csv(static_path) %>% semi_join(C_raw, by = "package") -> P

read_fst(calls_path) %>% as_tibble() -> E_raw

E_raw %>% select(
  run_package = package, #      The package which was run to trigger the eval
  ev_variant = eval_function, # Which eval variant was called
  duplicates = nb_ev_calls, #   Number of identical eval calls (weight)
  src_ref = eval_call_srcref, # Source ref for the eval call site
  file, #                       Script file name (name of this run)
  ev_package = eval_source, #   Package of the call site
  type = expr_resolved_type_tag, #   Argument type
  ev_expr = expr_expression,
  expr_resolved,
  func = expr_resolved_function, # the function (if any) at the root of the AST of expr_resolved
  parsed = expr_parsed_expression
) -> E

C_raw -> C
corpus_size <- nrow(C)
```

Some useful numbers:
```{r}
nb_call_sites <- E %>%
  pull(src_ref) %>%
  n_distinct()
```

# Most frequent expressions

Before resolving, so as the programmer wrote them:

```{r}
E %>% count(ev_expr, wt = duplicates, sort = TRUE)
```

After resolving:

```{r}
E %>% count(expr_resolved, wt = duplicates, sort = TRUE)
```

# Resolved expressions

Are expressions resolved to very diverse values or expressions?
If they are uniquely resolved and it is a value, then we can get rid of the `eval`.

```{r}

resolved_variety <- E %>%
  group_by(ev_expr, src_ref) %>%
  summarise(n = n_distinct(expr_resolved)) %>%
  arrange(desc(n))
resolved_variety %>% datatable()

nb_one_resolved <- resolved_variety %>%
  filter(n == 1) %>%
  nrow()
```

There are `r nb_one_resolved` call sites that are resolved to only one expression, out of `r nrow(resolved_variety)`.

We can now count how many of these call sites with only one resolved expression resolve to a value type.

```{r}
one_resolved_type <- E %>%
  group_by(src_ref) %>%
  mutate(n_resolved = n_distinct(expr_resolved)) %>%
  filter(n_resolved == 1) %>%
  ungroup() %>%
  select(src_ref, n_resolved, type) %>%
  distinct()

nb_one_resolved_type <- one_resolved_type %>% nrow()
```

`r nb_one_resolved_type` call sites resolved to one expression, of only one type.

```{r}
nb_value_type <- one_resolved_type %>% filter(type %in% c("character", "double", "integer", "logical", "NULL"))
```

There are `r nb_value_type %>% nrow()` such call sites.

# Expressions common to several call sites and packages

It emphasizes which expressions are a common pattern across various code bases.

```{r}
widespread_expr <- E %>%
  group_by(expr_resolved) %>%
  summarize(n_call_sites = n_distinct(src_ref), n_packages = n_distinct(ev_package))
```

```{r}
widespread_expr %>%
  arrange(desc(n_call_sites)) %>%
  datatable()
```

```{r}
widespread_expr %>%
  arrange(desc(n_packages)) %>%
  datatable()
```

For instance, for the most widespread expression, both in terms of call sites and packages, `mf`, we show here the packages sorted by decreasing number of call sites with this expression name.

```{r}
E %>%
  filter(ev_expr == "mf") %>%
  group_by(ev_package) %>%
  summarise(n = n_distinct(src_ref)) %>%
  arrange(desc(n))
```

Manual inspection seems to show that `mf` is often used as a result of a `match.call`:

- https://github.com/cran/PMCMRplus/blob/4f5234542aa0a60642b230bea7efb668fb16a140/R/mandelkTest.R#L121 
- https://github.com/fbertran/plsRglm/blob/dfa4e54ea02bca8bf04d29bb65dc7dba611927c9/R/plsRglmmodel.default.R#L4
- https://github.com/cran/RWeka/blob/55030d5513bf37760d1f6e2f1a49495977113276/R/classifiers.R#L42 with a comment _## The "usual" way of creating a model frame from the call._ (sic!)

This is the common pattern when a call is retrieved and then modified, then evaluated.

We can also look at the functions called in the expression:

```{r}
widespread_function <- E %>%
  filter(!is.na(func)) %>%
  group_by(func) %>%
  summarize(n_call_sites = n_distinct(src_ref), n_packages = n_distinct(ev_package))
```

# Expressions using `parse`, `str2lang`, and `str2expression`

```{r}
parsed_expr <- E %>% filter(!is.na(parsed))

nb_parsed_call_sites <- parsed_expr %>%
  pull(src_ref) %>%
  n_distinct()
```

There are `r nb_parsed_call_sites` call sites using those functions in their argument, out of the `r nb_call_sites` in total, i.e. `r ratio(nb_parsed_call_sites, nb_call_sites)`%.

# Arithmetic expressions

First, we only keep rows where the root is an arithmetic function, and the rows with a constant number

```{r}
arith_func <- c("/", "-", "*", "+", "^", "log", "sqrt", "exp", "max", "min", "cos", "sin", "abs", "atan", ":")

arithm <- E %>% filter(func %in% arith_func)
```

We want to keep only the ones that only use these functions and constants.

```{r}
is_arithm_expr <- function(exp) {
  if (is.call(exp)) {
    function_name <- exp[[1]]
    function_args <- exp[-1]
    return(as.character(function_name) %in% arith_func && every(function_args, is_arithm_expr))
  }
  else if (is.expression(exp)) {
    return(every(exp, is_arithm_expr))
  }
  else if (typeof(exp) %in% c("integer", "double", "complex")) {
    return(TRUE)
  }
  else {
    return(FALSE)
  }
}

arithm_folded <- arithm %>% filter(map_lgl(expr_resolved, function(exp) {
  ast <- NA
  try(ast <- parse(text = exp)[[1]], silent = TRUE)
  # some expr_resolved have been truncated so we mark them as FALSE (even though they could be true)
  if (length(ast) > 1 || !is.na(ast)) {
    return(is_arithm_expr(ast))
  }
  else {
    return(FALSE)
  }
}))
```

It corresponds to these call sites and packages:

```{r}
n_arithm_call_sites <- arithm_folded %>%
  pull(src_ref) %>%
  n_distinct()

arithm_folded %>%
  group_by(expr_resolved) %>%
  summarise(n_call_sites = n_distinct(src_ref), n_packages = n_distinct(ev_package)) %>%
  datatable()
```

The total number of call sites with those arithmetic on constants is `r n_arithm_call_sites`.

# Formulas

A formula is defined with the tilde `~` operator.

```{r}
with_formula <- E %>%
  filter(str_detect(expr_resolved, fixed("~")))
```

```{r}
n_with_formula <- with_formula %>%
  pull(src_ref) %>%
  n_distinct()
```

The total number of call sites with formulas is `r n_with_formula`.

```{r}
with_formula %>%
  group_by(expr_resolved) %>%
  summarise(n_call_sites = n_distinct(src_ref), n_packages = n_distinct(ev_package)) %>%
  datatable()
```

`~` is mostly used within a call to `model.frame`.

# Variables

Here, we analyze expressions with only a variable name.

```{r}
with_var <- E %>%
  filter(str_detect(expr_resolved, "^(?:\\.[:alpha]|[:alpha:])[[:alpha:][:digit:]_\\.]*$")) %>%
  filter(expr_resolved != "NA") # "NA" matches the regex of a variable name but is a value so we get rid of it
```

We can categorize them by number of call sites and number of packages:
```{r}
n_with_var <- with_var %>%
  pull(src_ref) %>%
  n_distinct()

with_var %>%
  group_by(expr_resolved) %>%
  summarise(n_call_sites = n_distinct(src_ref), n_packages = n_distinct(ev_package)) %>%
  datatable()
```

There are `r n_with_var` such call sites.

# Better normalization of resolved expressions

We fold all constant arithmetic expressions and replace all variable names by a placeholder.

```{r}
canonic_expr <- function(exp) {
  if (is.call(exp)) {
    function_name <- exp[[1]]
    function_args <- exp[-1]
    res <- map(function_args, canonic_expr)
    # Function_name can be larger than one is a function with a namespace stats::model or another higher-order function
    # For instance, (function(x, y) x + y)(3, 4)
    if (length(function_name) == 1 && as.character(function_name) %in% arith_func && every(res, function(v) {
      v == "NUM"
    })) {
      return("NUM")
    }
    else if (length(function_name) == 1 && function_name == "::") {
      str_c(as.character(function_args[[1]]), "::", as.character(function_args[[2]]))
    }
    else {
      function_name <- if (length(function_name) == 1) {
        as.character(function_name)
      }
      else {
        canonic_expr(function_name)
      }
      return(str_c(str_c(function_name, collapse = ", "), "(", str_c(res, collapse = ", "), ")"))
    }
  }
  else if (is.symbol(exp)) {
    return("VAR")
  }
  else if (is.expression(exp)) {
    return(str_c(map(exp, canonic_expr), collapse = ", "))
  }
  else if (typeof(exp) %in% c("integer", "double", "complex")) {
    return("NUM")
  }
  else {
    return(deparse1(exp))
  }
}

canonic_expr_str <- function(exp) {
  ast <- NA
  try(ast <- parse(text = exp)[[1]], silent = TRUE)
  # some expr_resolved have been truncated so we mark them as FALSE (even though they could be true)
  if (is.symbol(ast) || is.language(ast) || length(ast) > 1 || !is.na(ast)) {
    return(canonic_expr(ast))
  }
  else {
    return("CANONIZATION ERROR")
  }
}
```

For performance reason, we only apply this function to the unique resolved expressions.

```{r}
res_exprs <- E %>%
  select(expr_resolved) %>%
  unique()

res_exprs <- res_exprs %>% mutate(expr_canonic = map_chr(expr_resolved, canonic_expr_str))
```