### 2021-03-24 analysis/patterns2.Rmd
Patterns

---
title: "Patterns and anti-patterns"
output:
    html_document:
        gallery: false
        toc: true
        toc_depth: 3
        toc_float: true
        df_print: paged
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
params:
  base_dir: Data/package/
  calls_path: summarized.fst
  normalized_path: normalized-expressions.csv
  corpus_path: corpus.txt
  static_path: evals-static.csv
  force_rebuild: FALSE
---

```{r setup, include=FALSE}
library(tidyverse)
library(fst)
library(fs)
library(DT)
library(scales)
library(corrr)

knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) { # record the current time before each chunk
      now <<- Sys.time()
    } else { # calculate the time difference after a chunk
      res <- difftime(Sys.time(), now)
      # return a character string to show the time
      paste("Time for this code chunk to run:", round(res, 2), units(res))
    }
  }
}))
now <- Sys.time()
knitr::opts_chunk$set(
  echo = TRUE,
  fig.retina = 2,
  fig.width = 10,
  cache.lazy = FALSE,
  time_it = TRUE
)

source("inc/paths.R")
source("inc/latextags.R")
source("inc/functions.R")
theme_set(theme_minimal())

dataset_name <- basename(params$base_dir)

# Tags will be prefixed by the dataset name
create_tags(path(TAGS_DIR, paste0(dataset_name, "_patterns.tex")), prefix = dataset_name, default = TRUE)
# Also use ggsave_prefix instead of ggsave if you need to save a graph

calls_path <- paste0(params$base_dir, params$calls_path)
corpus_path <- paste0(params$base_dir, params$corpus_path)
static_path <- paste0(params$base_dir, params$static_path)
normalized_path <- paste0(params$base_dir, params$normalized_path)

lock_path <- paste0(params$base_dir, ".patterns-lock.fst")
min_path <- paste0(params$base_dir, "Epatterns.fst")
```

# Read

Read in the summarized calls data (`E_raw`) and the file that describes
which package belong to our corpus (`C_raw`). For conciseness, `E` is
the input data set with only the columns that we use. We also read the file with normalized expressions, `N`.

```{r read, message=TRUE, include=F}
tibble(package = read_lines(corpus_path)) -> C_raw
read_csv(static_path) %>% semi_join(C_raw, by = "package") -> P

rebuild <- TRUE

E_cano <- NULL

if (file.exists(lock_path)) {
  saved <- read.fst(lock_path)[[1,1]]
  if (saved == file.size(calls_path)) {
    read_fst(min_path) %>% as_tibble() -> E_cano
    rebuild <- FALSE
  }
}

if(params$force_rebuild || rebuild) {
  read_fst(calls_path) %>% as_tibble() -> E_raw

  read_csv(normalized_path, col_types = cols(
    minimized = col_character(),
    topcall = col_character(),
    is_model = col_logical(),
    has_fundef = col_logical(),
    has_calls = col_integer(),
    has_assigns = col_integer(),
    has_var = col_logical(),
    has_bracket = col_logical(),
    is_assign = col_logical(),
    is_value = col_logical(),
    is_ignore = col_logical(),
    has_dollar = col_logical(),
    has_user_call = col_logical(),
    has_block = col_logical(),
    has_meta_op = col_logical(),
    normalized = col_character(),
    trimmed = col_character(),
    hash = col_character()
  )) -> N

  E_raw %>% select(
    run_package = package, #      The package which was run to trigger the eval
    ev_variant = eval_function, # Which eval variant was called
    duplicates = nb_ev_calls, #   Number of identical eval calls (weight)
    src_ref = eval_call_srcref, # Source ref for the eval call site
    file, #                       Script file name (name of this run)
    ev_package = eval_source, #   Package of the call site
    type = expr_resolved_type_tag, #   Argument type
    ev_expr = expr_expression,
    expr_resolved,
    func = expr_resolved_function, # the function (if any) at the root of the AST of expr_resolved
    parsed = expr_parsed_expression,
    hash = expr_resolved_hash,
    match.call = expr_match_call,
    arg_size = expr_resolved_length, #   Size in characters of argument 
    ast_nodes = expr_resolved_nodes,
    n_op = interp_eval,
    envir = envir_expression,
    caller_srcref,
    caller_expression
  ) -> E

  E_cano <- E %>% left_join(N, by = "hash") # Common column is the hash of the full resolved expression

  E_cano %>% write_fst(min_path)
  sz <- file.size(calls_path) %>% as_tibble()
  write.fst(sz, lock_path)
}

C_raw -> C
corpus_size <- nrow(C)
```

Some useful numbers:
```{r}
nb_call_sites <- E_cano %>%
  pull(src_ref) %>%
  n_distinct()

r("Nb call sites", nb_call_sites)
```

There are `r nb_call_sites` call sites in the dataset.

# Variable lookup in another environment

Variables lookup are characterized by:

- symbol as the type of the expression
- a non-default environment

```{r}
variable_lookup <- E_cano %>% filter(type == "symbol", !is.na(envir))

nb_variable_lookup <- variable_lookup %>% pull(src_ref) %>% n_distinct()
```

There are `r nb_variable_lookup` sites with variable lookup, i.e `r ratio(nb_variable_lookup, nb_call_sites)`%.

# `match.call`

The goal is to get a call and modify it or use it.

```{r}
match.call_df <- E_cano %>% filter(!is.na(match.call))
nb_m.c_call_sites <- match.call_df %>% pull(src_ref) %>% n_distinct()

```

There are `r nb_m.c_call_sites` sites using `match.call`, i.e. `r ratio(nb_m.c_call_sites, nb_call_sites)`%.

```{r}
match.call_df %>% 
  group_by(topcall) %>%
  summarize(n_call_sites = n_distinct(src_ref)) %>% 
  arrange(desc(n_call_sites))
```

## Evaluating some of the call arguments

In that case, `topcall` is `NA`, `$` or `[`.

```{r}
mc_arg_functions <- c(NA_character_, "$", "[")

nb_arg_mc_sites <- match.call_df %>% 
  filter(topcall %in% mc_arg_functions) %>% 
  pull(src_ref) %>% 
  n_distinct()
```

it amounts for `r nb_arg_mc_sites` call sites, i.e `r ratio(nb_arg_mc_sites, nb_m.c_call_sites)`%.

## `match.call` and statistical models

Veyr often, `match.call` is used to build statistical models, by replacing the call function by `model.frame` (but also other statistical models, such as `glm`).

```{r}
stat_functions <- c("model.frame", "glm", "lm", "plm", "elo.model.frame", "lmer", "randomForest", "betareg", "nlreg", "model.frame.default", "model.matrix", "arima", "glm.nb", "glmer") # There are many more

nb_stat_call_sites <- match.call_df %>% 
  filter(topcall %in% stat_functions) %>% 
  pull(src_ref) %>% 
  n_distinct()

```

it amounts for `r nb_stat_call_sites` call sites, i.e `r ratio(nb_stat_call_sites, nb_m.c_call_sites)`%.

# Parsing strings

As a first approximation, it corresponds to type `expression` (although `expr` can also build such expressions). However, it is an under-approximation, because the expression could be later used...

```{r}
parse_eval <- E_cano %>% filter(type == "expressions" | str_detect(ev_expr, fixed("parse")))

nb_parse_eval <- parse_eval %>% pull(src_ref) %>% n_distinct()
```

There are `r nb_parse_eval` sites using `parse`, i.e. `r ratio(nb_parse_eval, nb_call_sites)`%.

## Logging

https://github.com/cran/Rcmdr/blob/4b196764a44a047f8c7be3f4c629f52bee2c0c80/R/commander.R#L1301

The goal is to print the command and then to execute it.

## Formulas

```{r}
formulas <- E_cano %>% filter(str_detect(ev_expr, fixed("parse")), topcall == "~" | str_detect(ev_expr, fixed("formula")))

nb_formula_sites <- formulas %>% pull(src_ref) %>% n_distinct()
```

There are `r nb_formula_sites` sites building formulas with `parse`, i.e. `r ratio(nb_formula_sites, nb_call_sites)`%.

# `Parse` and plotting functions

We approximate the plotting function by just looking for `plot` in their name:

```{r}
plotting <- E_cano %>% filter(str_detect(normalized, fixed("plot"))) # not in topcall because f it results from parse...

nb_plotting_call_sites <- plotting %>% pull(src_ref) %>% n_distinct()
```

There are `r nb_plotting_call_sites` plotting call sites, i.e. `r ratio(nb_plotting_call_sites, nb_call_sites)`% and `r ratio(nb_plotting_call_sites, nb_parse_eval)`% of the parsing sites.

Are they built using parse? For that, we look at the non-resolved expression (`parse` will disappear after).

Currently, we use a regex. When `parsed` column is fixed, we will get more accurate results.

```{r}
parse_plot <- plotting %>% filter(str_detect(ev_expr, fixed("parse")))

nb_parse_plot_call_sites <- parse_plot %>% pull(src_ref) %>% n_distinct()
```

This already amounts to `r nb_parse_plot_call_sites` call sites, i.e. `r ratio(nb_parse_plot_call_sites, nb_plotting_call_sites)`% of the plotting call sites.

# Complex blocks

`eval` can evaluate sometimes complex blocks.

```{r}
blocks <- E_cano %>% filter(has_block)

nb_blocks_call_sites <- blocks %>% pull(src_ref) %>% n_distinct()
```

There are `r nb_blocks_call_sites` such call sites, i.e. `r ratio(nb_blocks_call_sites, nb_call_sites)`%.

`eval` is used to insert blocks of code inside the function, for instance as a preprocessing and postprocessing step. There are some examples in the VGAM package. In that case, `{` will be at the root of the AST.

```{r}
pre_post <- blocks %>% filter(topcall == "{")

nb_pre_post_call_sites <- pre_post %>% pull(src_ref) %>% n_distinct()
```

It happens in `r nb_pre_post_call_sites` call sites so most of the call sites with blocks (`r ratio(nb_pre_post_call_sites, nb_blocks_call_sites)`%).

# Symbolic computations

`D` operator to derive an expression symbolically, `Simplify`

For instance: https://github.com/cran/Deriv/blob/9ca174a959d8eb39b0a1f1a70de26d147a837176/R/Simplify.R#L572

Usually on an expression object, not a language object. Environment type seems to be `list` in that case, most often.

# Infix operators

```{r}
infix <- E_cano %>% filter(str_detect(topcall, "%[^%]*%"))

nb_infix_call_sites <- infix %>% pull(src_ref) %>% n_distinct()
```

It happens in `r nb_infix_call_sites` call sites, i.e. `r ratio(nb_infix_call_sites, nb_call_sites)`%.

The packages that use or define those infix operators:

```{r}
infix %>% group_by(ev_package) %>% summarize(n_call_sites = n_distinct(src_ref)) %>% arrange(desc(n_call_sites)) %>% datatable()
```

# Meta-programming

As a first approximation, it corresponds to type `language`.

```{r}
parse_eval <- E_cano %>% filter(type == "language")

nb_parse_eval <- parse_eval %>% pull(src_ref) %>% n_distinct()
```

## selecting columns in a dataframe

https://github.com/cran/epiDisplay/blob/d3d129a690a84b95c9aec437690227f1d50e3394/R/epiDisplay.R#L5915

```r 
 selected <- eval(substitute(vars), nl, parent.frame())
```

`vars` correspond to the columns in the dataframe we want to use.

Some examples on how it is called. Columns can be selected by name or by column number.

```r 
tableStack(bakedham:fruitsalad, .data)

tableStack(vars=4:25, .data, by=Origin)

tableStack(c(qa1,qa13:qa18,mean.score,total.score), .data, by=sex, test=FALSE)
```

# Non-necessary `eval`

## Values into `eval`