### 2021-04-05 analysis/corpus.Rmd
Add traced evals figure

---
title: "Corpus"
output:
    html_document:
        gallery: false
        toc: true
        toc_depth: 3
        toc_float: true
        df_print: paged
editor_options: 
  chunk_output_type: console
params:
  base_dir: Data/
---

```{r setup, include=FALSE}
local(for (x in c(
  "dplyr", 
  "DT",
  "fs",
  "fst",
  "ggplot2",
  "ggthemes",
  "lubridate",
  "readr",
  "purrr",
  "stringr",
  "tidyr",
  "viridis"
  )) {
  suppressPackageStartupMessages(library(x, character.only=TRUE))
})

knitr::opts_chunk$set(
  echo = TRUE,
  fig.align = "center",
  fig.retina = 2,
  fig.width = 10
)

source("inc/functions.R")
source("inc/latextags.R")
source("inc/paths.R")

create_tags(path(TAGS_DIR, "corpus.tex"), prefix = "", default = TRUE)
theme_set(theme_minimal())
```

## Load data

```{r load corpus}
PKG_S <- 
  read_fst(PACKAGE_SUM_CALLS_FILE) %>%
  # remove evals that are inside the scripts,
  # for PKG_C we are only interested in evals in packages
  filter(!str_starts(eval_call_srcref, "::global::main")) %>% 
  as_tibble()

PKG_evals_per_package <- 
  count(PKG_S, caller_package, eval_call_srcref) %>%
  count(caller_package) %>% 
  rename(
    package=caller_package, 
    evals_hit=n
  )

CRAN_C <- 
  read_fst(PACKAGE_CORPUS_FILE) %>%
  mutate_if(is.numeric, replace_na, 0) %>%
  mutate(
    evals=as.integer(evals), 
    package_code=package_native_code + package_r_code
  ) %>%
  filter(loadable) %>% 
  as_tibble()

PKG_C <- 
  CRAN_C %>%
  filter(in_corpus, package_code > 0) %>%
  left_join(PKG_evals_per_package, by="package") %>% 
  mutate(
    evals_hit=as.integer(evals_hit),
    # this is the case when there is eval in package loading code
    # this we will always run
    evals_hit=if_else(evals_hit > evals, evals, evals_hit),
    evals_coverage=evals_hit/evals
  ) %>% 
  mutate_if(is.numeric, replace_na, 0) %>%
  as_tibble()

CRAN_E <- 
  read_csv(PACKAGE_EVALS_STATIC_FILE) %>% 
  # TODO: There are very few NA, plus when srcref is NA, the rest of the row in NA
  # we should look at this
  filter(!is.na(srcref))

PKG_E <- semi_join(CRAN_E, PKG_C, by="package")

KAGGLE_ALL <- local({
  kernels <- read_csv(KAGGLE_KERNEL_FILE)
  evals <- read_csv(KAGGLE_EVALS_STATIC_FILE) %>% count(package) %>% rename(evals=n)
  left_join(kernels, evals, by="package") 
}) %>%
  mutate_if(is.numeric, replace_na, 0) %>% 
  mutate(evals=as.integer(evals)) %>% 
  as_tibble()

KAGGLE <- filter(KAGGLE_ALL, evals > 0)

PKG_undefined <- 
  read_fst(PACKAGE_UNDEFINED_FILE) %>% 
  as_tibble()

TRACE_LOG <- read_csv(PACKAGE_TRACE_LOG_FILE, col_types = cols(
  Seq = col_double(),
  Host = col_character(),
  Starttime = col_double(),
  JobRuntime = col_double(),
  Send = col_double(),
  Receive = col_double(),
  Exitval = col_double(),
  Signal = col_double(),
  Command = col_character(),
  V1 = col_character(),
  Stdout = col_logical(),
  Stderr = col_logical()
))

BASE_packages <- read_lines(BASE_CORPUS_FILE)
BASE_evals_per_package <- read_csv(BASE_EVALS_STATIC_FILE, col_types = cols(
  package = col_character(),
  fun_name = col_character(),
  srcref = col_character(),
  call_fun_name = col_character(),
  args = col_character()
))

BASE_C <- 
  tibble(package=BASE_packages) %>%
  left_join(rename(count(BASE_evals_per_package, package), evals=n), by="package") %>%
  left_join(rename(count(BASE_evals_per_package, package, fun_name) %>% count(package), funs_with_eval=n), by="package")
```

## Overview

```{r}
overview_table(
  r("base packages", BASE_C),
  r("base eval call sites", sum(BASE_C$evals)),
  r("base funs with evals", sum(BASE_C$funs_with_eval)),
)
```

### Packages

```{r}
cran_funs          <- sum(CRAN_C$funs_public) + sum(CRAN_C$funs_private)
cran_code          <- sum(CRAN_C$package_code)
cran_r_code        <- sum(CRAN_C$package_r_code)
cran_native_code   <- sum(CRAN_C$package_native_code)
cran_runnable_code <- sum(CRAN_C$runnable_code)

pkg_funs           <- sum(PKG_C$funs_public) + sum(PKG_C$funs_private)
pkg_funs_with_eval <- sum(PKG_C$funs_with_eval)
pkg_code           <- sum(PKG_C$package_code)
pkg_r_code         <- sum(PKG_C$package_r_code)
pkg_native_code    <- sum(PKG_C$package_native_code)
pkg_evals          <- PKG_C$evals
pkg_evals_hit      <- PKG_C$evals_hit
pkg_runnable_code  <- sum(PKG_C$runnable_code)
pkg_eval_calls     <- sum(PKG_S$nb_ev_calls)

pkg_failed_programs <- filter(TRACE_LOG, Exitval > 1)

overview_table(
  r("CRAN available packages", 15962), #length(available.packages()[,1])), # from the PKG_C mirror
  r("CRAN packages", CRAN_C),
  r("PKG packages", PKG_C),
  r("PKG packages ratio", ratio(PKG_C, CRAN_C)),
  r("CRAN code", cran_code),
  r("CRAN r code", sum(CRAN_C$package_r_code)),
  r("CRAN native code", sum(CRAN_C$package_native_code)),
  r("PKG code", pkg_code),
  r("PKG code ratio", ratio(pkg_code, cran_code)),
  r("PKG r code", pkg_r_code),
  r("PKG r code ratio", ratio(pkg_r_code, cran_r_code)),
  r("PKG native code", pkg_native_code),
  r("PKG native code ratio", ratio(pkg_native_code, cran_native_code)),
  r("CRAN funs", cran_funs),
  r("PKG funs", pkg_funs),
  r("PKG funs with eval", pkg_funs_with_eval),
  r("PKG funs with eval ratio", ratio(pkg_funs_with_eval, pkg_funs)),
  r("CRAN funs with eval ratio", ratio(pkg_funs_with_eval, cran_funs)),
  r("PKG eval call sites", sum(pkg_evals)),
  r("PKG hit eval call sites", sum(pkg_evals_hit)),
  r("PKG hit eval call sites avg ratio", percent(mean(pkg_evals_hit/pkg_evals))),
  r("PKG code coverage", percent(mean(PKG_C$coverage)/100)),
  r("CRAN code coverage", percent(mean(CRAN_C$coverage)/100)),
  r("CRAN runnable scripts", sum(CRAN_C$runnable_files)),
  r("CRAN runnable code", cran_runnable_code),
  r("CRAN runnable code examples", sum(CRAN_C$runnable_code_examples)),
  r("CRAN runnable code tests", sum(CRAN_C$runnable_code_tests)),
  r("CRAN runnable code vignettes", sum(CRAN_C$runnable_code_vignettes)),
  r("PKG undefined", PKG_undefined),
  r("PKG undefined ratio", ratio(PKG_undefined, pkg_eval_calls)),
  r("PKG failed programs", pkg_failed_programs),
  r("PKG failed programs ratio", ratio(pkg_failed_programs, TRACE_LOG))
)
```

### Kaggle

```{r}
kaggle <- distinct(KAGGLE_ALL, hash, .keep_all=TRUE)
kaggle_code <- sum(kaggle$code)
kaggle_evals <- filter(kaggle, evals > 0)
kaggle_evals_code <- sum(kaggle_evals$code)

overview_table(
  r("Kaggle kernels", KAGGLE_ALL),
  r("Kaggle duplicates", nrow(KAGGLE_ALL)-nrow(kaggle)),
  r("Kaggle unique", kaggle),
  r("Kaggle code", kaggle_code),
  r("Kaggle competitions", count(kaggle, competition)),
  r("Kaggle with evals", kaggle_evals),
  r("Kaggle with evals code", kaggle_evals_code),
  r("Kaggle with evals code ratio", ratio(kaggle_evals_code, kaggle_code))
)
```

```{r}
cs_count(PKG_C, evals) %>% print(n=Inf)
```

### Details

```{r}
max_eval_cs_idx <- which.max(PKG_C$evals)
max_eval_cs_package <- PKG_C$package[max_eval_cs_idx]
max_eval_cs_count <- PKG_C$evals[max_eval_cs_idx]

overview_table(
  r("PKG eval call sites median", median(PKG_C$evals)),
  r("Max eval call sites count", max_eval_cs_count),
  r("Max eval call sites package", max_eval_cs_package),
  r("Max eval call sites rest", round(max(PKG_C$evals[-max_eval_cs_idx])/100)*100)
)
```

### All

```{r}
overview_table(
  r("Corpus code", kaggle_code + pkg_runnable_code)
)
```

### Matching evals

```{r}
matched_eval_callsites <-
  left_join(
    CRAN_eval_callsites, 
    count(CRAN_eval_traces, eval_call_srcref), 
    by=c("srcref"="eval_call_srcref")
  ) %>%
  mutate(
    n=replace_na(n, 0)
  ) %>%
  rename(hit=n)
```

The dynamic analysis traced `r fmt(ratio(filter(matched_eval_callsites, hit>0), matched_eval_callsites))`

### Plot

```{r}
eval_call_sites_hist <- function(df, df_all, y_lab, max_evals=20L) {
  median_evals <- median(df$evals)
  df %>% 
    mutate(
      evals=ifelse(evals < max_evals, evals, max_evals)
    ) %>%
    ggplot(aes(x=evals)) + 
    geom_histogram(binwidth=1, color="black", fill="blue", alpha=0.3) +
    geom_vline(aes(xintercept=median_evals), linetype="dashed", color="red", size=0.2) +
    scale_x_continuous(
      limits=c(0, max_evals + 1), 
      breaks=as.integer(sort(c(seq(1, max_evals, length.out=4), median_evals))), 
      labels=function(x) c(head(x, -1), str_c(tail(x, 1), "+"))
    ) +
    scale_y_continuous(
      labels=scales::comma_format(accuracy=1),
      sec.axis=sec_axis(
        trans=~./nrow(df_all), 
        labels=scales::percent_format(accuracy=.1), 
        name=y_lab)
    ) +
    labs(y=y_lab) +
    theme(text = element_text(size=21),
          axis.title.x = element_blank(),
          panel.grid.minor.x = element_blank(), aspect.ratio = 4/3) 
}

eval_call_sites_hist(PKG_C, CRAN_C, "CRAN packages")
ggsave(path(PLOT_DIR, "pkgs-eval-callsites-hist.pdf"))
eval_call_sites_hist(KAGGLE, KAGGLE_ALL, "KAGGLE programs")
ggsave(path(PLOT_DIR, "kaggle-eval-callsites-hist.pdf"))
```

### Eval plot

```{r}
eval_plot_df <- PKG_C %>%
  filter(package_r_code > 0) %>%
  select(package, evals, evals_hit, package_r_code)

size_breaks <- c(1e3, 5e3, 10e3, 25e3, 50e3, 100e3)
# size_labels <- str_c(
#   c("≤ ", rep("> ", 6)),
#   scales::comma(accuracy = 1, x=size_breaks[-1])
# )

eval_plot_df %>% 
#  filter(evals > 0, evals_hit > 0) %>%
  mutate(
    # size=cut(package_r_code, size_breaks, size_labels, include.lowest=T),
    label=ifelse(evals %in% tail(sort(evals),3), package, NA)
  ) %>%
  ggplot(aes(x=evals, y=evals_hit, size=package_r_code, label=label)) + 
    geom_jitter(alpha=0.3) +
    geom_text(nudge_y = .15, check_overlap = TRUE, size=4) +
    scale_x_continuous(
      trans="log10",
      breaks=function(x) {
        f <- scales::breaks_log()
        sort(c(f(x), max(eval_plot_df$evals)))
      },
      labels=scales::comma_format(accuracy = 1)
    ) +
    scale_y_continuous(
      trans="pseudo_log",
      breaks=function(x) {
        f <- scales::breaks_log()
        sort(c(0, 1, f(c(1, x[2])), max(eval_plot_df$evals_hit)))
      },
      labels=scales::comma_format(accuracy = 1)
    ) +
    scale_size_binned(
      range=c(1,10),
      breaks = size_breaks,
      labels=scales::comma_format(accuracy = 1)
    ) +
#    geom_smooth(method = "lm", se = FALSE, show.legend = FALSE, linetype="dashed", color="red", size=0.2) +
    labs(x="Eval call sites (log)", y="Traced eval call sites (log)", size="Package code size") +
    theme(legend.position = c(0.15, 0.8))
ggsave(path(PLOT_DIR, "traced-eval-callsites.pdf"))
```