### 2021-04-12 analysis/side-effects.Rmd

---
title: "Side-effects"
output:
    html_document:
        gallery: false
        toc: true
        toc_depth: 3
        toc_float: true
        df_print: paged
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
params:
  base_dir: Data
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(DT)
library(digest)
library(fs)
library(fst)
library(purrr)
library(tidyverse)

source("inc/latextags.R")
source("inc/functions.R")
source("inc/paths.R")

create_tags(path(TAGS_DIR, "side-effects.tex"), prefix = "", default = TRUE)
```

```{r load data}
SE <- read_fst(PACKAGE_SIDE_EFFECTS_FILE) %>% as_tibble
```

```{r fix withr::defer callsite}

```

```{r fix S3 methods}

```

```{r}
cs_count(SE, eval_call_srcref, wt=N)
cs_count(SE, variable, wt=N)
```

```{r}
overview_table(
  r("Number of side effects", sum(SE$N))
)
```

## Unknwon srcref

```{r}
unknown_srcref <- filter(SE, eval_call_srcref=="???")
```

## Where does the side effects happen

- in the caller - source is `call:fun` where `fun` is the `calle_function` 
  - most of this should be class that do not specify any parameter to `envir`
- in the supplied `envir` parameter -- `in_envir` is true
  - in this case the details are in the `environment_class`
- in a package - the source is `pkg:`
- in the `.userHookEnv` - the source is `userHookEnv`
- in another call
- is S4, R6

```{r}

```