### 2021-04-16 analysis/side-effects.Rmd

---
title: "Side-effects"
output:
    html_document:
        gallery: false
        toc: true
        toc_depth: 3
        toc_float: true
        df_print: paged
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
params:
  base_dir: Data
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(DT)
library(digest)
library(fs)
library(fst)
library(purrr)
library(tidyverse)
library(viridis)
library(xtable)

source("inc/latextags.R")
source("inc/functions.R")
source("inc/paths.R")

create_tags(path(TAGS_DIR, "side-effects.tex"), prefix = "SE", default = TRUE)
```

```{r load data}
SE_all <- read_fst(PACKAGE_SIDE_EFFECTS_FILE) %>% 
  as_tibble %>%
  filter(transitive==0)
```

```{r missing calls}
SE_missing_eval_call <- filter(SE_all, is.na(eval_function))
SE_all <- filter(SE_all, !is.na(eval_function))
```

```{r deal with empty srcrefs}
SE <- mutate(SE_all, eval_call_srcref=case_when(
  eval_call_srcref == "???" & expr_expression == "handler$expr" & envir_expression == "handler$envir" ~ "::withr::execute_handlers::1",
  eval_call_srcref == "???" & expr_expression == "quote({\n    old.options <- options(dplyr.summarise.inform = FALSE)\n    o" ~ "::dplyr::tally::1",
  eval_call_srcref == "???" & expr_expression == "quote({\n    na.action <- attr(mf, \"na.action\")\n    why_omit <- attr(na.a" ~ "::estimatr::cleam_mode_data::1",
  TRUE ~ eval_call_srcref
))

SE_missing_sites <- filter(SE, eval_call_srcref=="???")
SE <- filter(SE, eval_call_srcref!="???")
```

```{r remove Random.seed}
SE_random_seed <- filter(SE, variable == ".Random.seed", !str_detect(expr_resolved, "set.seed"))
SE <- filter(SE, variable != ".Random.seed" | str_detect(expr_resolved, "set.seed"))
```

```{r remove internal}
SE_internal <- filter(SE, str_starts(source, fixed("internal:")))
SE <- filter(SE, !str_starts(source, fixed("internal:")))
```

```{r load static evals}
ES <- read_csv(PACKAGE_EVALS_STATIC_FILE) %>% rename(eval_call_srcref=srcref)
```

```{r load summarize data}
S <- read_fst(PACKAGE_SUM_CALLS_FILE) %>% as_tibble
```

```{r}
CRAN <- read_fst(PACKAGE_CORPUS_FILE) %>% as_tibble
PKG <- filter(CRAN, in_corpus) 
```

TODO: fix the srcrefs and caller packages / functions

```{r}
SE <- mutate(
  SE,
  caller=str_replace(eval_call_srcref, "::(.*)::(.*)::.*", "\\1::\\2")
)

S <- mutate(
  S,
  caller=str_replace(eval_call_srcref, "::(.*)::(.*)::.*", "\\1::\\2"),
  caller_package=str_replace(eval_call_srcref, "::(.*)::.*::.*", "\\1")
)
```

## Overview

```{r}
S_calls <- sum(S$nb_ev_calls)
S_sites <- count(S, eval_call_srcref)
S_pkgs <- count(S, eval_source)
S_funs <- count(S, caller)

SE_all_N <- sum(SE_all$N)
SE_all_sites <- count(SE_all, eval_call_srcref)
SE_all_calls <- count(SE_all, cid)

SE_internal_N <- sum(SE_internal$N)
SE_internal_sites <- count(SE_internal, eval_call_srcref)

SE_random_seed_N <- sum(SE_random_seed$N)
SE_random_seed_sites <- count(SE_random_seed, eval_call_srcref)

SE_N <- sum(SE$N)
SE_calls <- count(SE, cid)
SE_sites <- count(SE, eval_call_srcref)
SE_funs <- count(SE, caller)
SE_pkgs <- count(SE, caller_package)
SE_missing_N <- sum(SE_missing_sites$N)
```

```{r}
overview_table(
  r("All", SE_all_N),
  r("All calls", SE_all_calls),
  r("All sites", SE_all_sites),

  r("Internal", SE_internal_N),
  r("Internal ratio", ratio(SE_internal_N, SE_all_N)),

  r("Internal sites", SE_internal_sites),
  r("Internal sites ratio", ratio(SE_internal_sites, SE_all_sites)),

  r("Random seed", SE_random_seed_N),
  r("Random seed ratio", ratio(SE_random_seed_N, SE_all_N)),

  r("Random seed sites", SE_random_seed_sites),
  r("Random seed sites ratio", ratio(count(SE_random_seed, eval_call_srcref), SE_all_sites)),

  r("Missing SE sites", SE_missing_N),
  r("Missing SE sites ratio", ratio(SE_missing_N, SE_N)),

  r("User", SE_N),
  r("User ratio", ratio(SE_N, SE_all_N)),

  r("User calls", SE_calls),
  r("User calls ratio", ratio(SE_calls, SE_all_calls)),
  r("User calls to all ratio", ratio(SE_calls, S_calls)),

  r("User sites", SE_sites),
  r("User sites ratio", ratio(SE_sites, SE_all_sites)),
  r("User sites to all ratio", ratio(SE_sites, S_sites)),

  r("User functions", SE_funs),
  r("User functions to all ratio", ratio(SE_funs, S_funs)),
  r("User packages", SE_pkgs),
  r("User packages to all ratio", ratio(SE_pkgs, S_pkgs))
)
```

## Classification

```{r functions}
SE <- mutate(
  SE, 
  class=case_when(
    source == str_c("call:", caller_function)           ~ "Local",
    str_starts(source, fixed("call:"))                  ~ "Function",
    str_detect(source, "^R6:.*R6.*") | source=="???S4"  ~ "Object",
    source == "pkg:global"                              ~ "Global",
    str_starts(source, fixed("pkg:"))                   ~ "Package",
    TRUE                                                ~ "Synthetic"
  )
)
```

```{r}
sites_classes <- 
  SE %>%
  group_by(eval_call_srcref) %>%
  summarise(n_class=n_distinct(class), classes=str_c(sort(unique(class)), collapse=", "))

cs_count(sites_classes, n_class)
cs_count(sites_classes, classes)
cs_count(SE, class)
```

```{r}
funs <- cs_count(SE, caller)
```

```{r}
sites_envir <- 
  SE %>%
  group_by(eval_call_srcref) %>% 
  summarize(
    in_envir=case_when(
      all(!in_envir) ~ 0,
      all(in_envir)  ~ 1,
      TRUE ~ 2
    )
  ) %>% 
  ungroup()
```

```{r}
sites_not_in_envir <- filter(sites_envir, in_envir==0)
sites_in_envir <- filter(sites_envir, in_envir==1)
sites_both_envir <- filter(sites_envir, in_envir==2)

overview_table(
  r("Funs nighty", filter(funs, cp <= 90)),
  r("Funs ten", filter(funs, cp > 90)),
  r("Sites not in envir", sites_not_in_envir),
  r("Sites not in envir ratio", ratio(sites_not_in_envir, sites_envir)),
  r("Sites in envir", sites_in_envir),
  r("Sites in envir ratio", ratio(sites_in_envir, sites_envir)),
  r("Sites both envir", sites_both_envir),
  r("Sites both envir ratio", ratio(sites_both_envir, sites_envir))
)
```

```{r}
sum_sites_classes <- 
  sites_classes %>%
  mutate(
    classes=if_else(str_detect(classes, "[,;]"), "multiple", classes)
  ) %>%
  cs_count(classes) %>%
  mutate(
    p=fmt(p, suffix="\\%", digit=0),
    o=if_else(str_starts(classes, fixed("\\")), 0L, n)
  ) %>% 
  arrange(desc(o)) %>%
  select(-cp, -o)

# sum_sites_classes %>%
#   mutate(classes=if_else(classes=="multiple", "\\emph{multiple}", classes)) %>%
#   xtable(
#     align="llrr",
#   ) %>%
#   print(
#     file=file.path(TAGS_DIR, "table-top-sites-classes-se.tex"),
#     comment=FALSE, 
#     include.colnames=FALSE,
#     include.rownames=FALSE,
#     only.contents=TRUE,
#     sanitize.text.function=identity,
#     timestamp=FALSE
#  )
```

```{r}
sites_class_n <- cs_count(sites_classes, n_class)

overview_table(
  r("Sites in one class", percent(filter(sites_class_n, n_class==1)$p/100))
)
```

## Function classification

```{r}
funs_srcrefs <- 
  left_join(
    mutate(ES, fun=str_c(package, "::", fun_name)), 
    transmute(count(S, eval_call_srcref), eval_call_srcref, hit=1), 
    by="eval_call_srcref"
  ) %>%
  mutate(hit=replace_na(hit, 0)) %>%
  select(-args, -call_fun_name) %>%
  left_join(
    select(sites_classes, eval_call_srcref, classes),
    by="eval_call_srcref"
  )

funs_cov <-
  funs_srcrefs %>% 
  group_by(fun) %>%
  summarize(cov=sum(hit)/n())

funs_classes <- 
#  semi_join(funs_srcrefs, filter(funs_cov, cov==1), by="fun") %>%
  semi_join(funs_srcrefs, funs_cov, by="fun") %>%
  group_by(fun) %>%
  summarize(
    cov=sum(hit)/n(),
    classes=str_c(unique(classes[!is.na(classes)]), collapse=";")
  )

cs_count(funs_classes, cov)
cs_count(funs_classes, classes)
```

```{r}
sum_funs_classes <-
  funs_classes %>%
    filter(cov==1, classes != "") %>%
    mutate(
      classes=if_else(str_detect(classes, "[,;]"), "multiple", classes)
    ) %>%
    cs_count(classes) %>%
    mutate(
      p=fmt(p, suffix="\\%", digit=0),
      o=if_else(str_starts(classes, fixed("\\")), 0L, n)
    ) %>% 
    arrange(desc(o)) %>%
    select(-cp, -o)

# sum_funs_classes %>%
#   mutate(classes=if_else(classes=="multiple", "\\emph{multiple}", classes)) %>%
#   xtable(
#     align="llrr",
#   ) %>% 
#   print(
#     file=file.path(TAGS_DIR, "table-top-sites-classes-funs.tex"),
#     comment=FALSE, 
#     include.colnames=FALSE,
#     include.rownames=FALSE,
#     only.contents=TRUE,
#     sanitize.text.function=identity,
#     timestamp=FALSE
#   )
```

```{r}
left_join(
  rename(sum_sites_classes, site_n=n, site_p=p),
  rename(sum_funs_classes, fun_n=n, fun_p=p)
) %>%
  mutate(
    o=if_else(classes=="multiple", 0L, site_n),
    classes=if_else(classes=="multiple", "\\emph{multiple}", classes),
    fun_n=replace_na(fun_n, "0"),
    fun_p=replace_na(fun_p, "0\\%"),
  ) %>%
  arrange(desc(o)) %>%
  select(-o) %>%
  xtable(
    align="llrrrr"
  ) %>%
  print(
    file=file.path(TAGS_DIR, "table-se-target-envs.tex"),
    comment=FALSE, 
    include.colnames=FALSE,
    include.rownames=FALSE,
    only.contents=TRUE,
    sanitize.text.function=identity,
    timestamp=FALSE
  )
```

```{r}
event <- cs_count(SE, event, wt=N)
event_calls <- cs_count(SE, event)
event_site <- cs_count(SE, eval_call_srcref, event) %>% cs_count(event)

left_join(
  select(event, event, n_event=n, p_event=p),
  select(event_calls, event, n_call=n, p_call=p),
  by="event"
) %>%
  left_join(
    select(event_site, event, n_site=n, p_site=p),
    by="event"
  ) %>% 
  mutate(event=case_when(
    event=="def" ~ "definition",
    event=="asn" ~ "update",
    event=="rvl" ~ "removal"
  )) %>%
  mutate_if(is.double, ~fmt(., suffix="%", digit=0)) %>%
  mutate_if(is.integer, ~fmt(.)) %>%
  xtable(
    align="llrrrrrr"
  ) %>%
  print(
    file=file.path(TAGS_DIR, "table-se-types.tex"),
    comment=FALSE, 
    include.colnames=FALSE,
    include.rownames=FALSE,
    only.contents=TRUE,
    timestamp=FALSE
  )
```

```{r}
event_rvl_class <- filter(SE, event=="rvl") %>% cs_count(class)

overview_table(
  r("event rem from synthetic", percent(filter(event_rvl_class, class=="Synthetic")$p/100)),
  r("event rem from function", percent(filter(event_rvl_class, class=="Function")$p/100)),
  r("event rem from local", percent(filter(event_rvl_class, class=="Local")$p/100))
)
```

## Types

```{r}
SEXP_TYPES <- tribble(
  ~type, ~type_name,
  0, "NILSXP",
  1, "SYMSXP",
  2, "LISTSXP",
  3, "CLOSXP",
  4, "ENVSXP",
  5, "PROMSXP",
  6, "LANGSXP",
  7, "SPECIALSXP",
  8, "BUILTINSXP",
  9, "CHARSXP",
  10, "LGLSXP",
  13, "INTSXP",
  14, "REALSXP",
  15, "CPLXSXP",
  16, "STRSXP",
  17, "DOTSXP",
  18, "ANYSXP",
  19, "VECSXP",
  20, "EXPRSXP",
  21, "BCODESXP",
  22, "EXTPTRSXP",
  23, "WEAKREFSXP",
  24, "RAWSXP",
  25, "S4SXP",
  30, "NEWSXP",
  31, "FREESXP",
  99, "FUNSXP"
) %>%
  mutate(
    type_name=str_replace(type_name, "SXP", "")
  )
```

```{r}
SE1 <- left_join(SE, SEXP_TYPES, by="type") %>%
  select(-type) %>%
  rename(type=type_name) %>%
  filter(event != "rvl") %>%
  mutate(
    stype=case_when(
      type == "CLO" ~ "Closure",
      type %in% c("STR","INT","LGL","RAW","REAL", "CPLX") ~ "Basic",
      type %in% c("ENV") ~ "Environment",
      type %in% c("VEC") ~ "List",
      type %in% c("SYM","LANG","EXPR") ~ "Expression",
      type %in% c("NIL") ~ "NULL",
      type %in% c("S4") ~ "S4",
      TRUE ~ "Other"
    )
  )

cs_count(SE1, stype)

class_type <- 
  SE1 %>%
  filter(event=="asn") %>%
  count(eval_call_srcref, stype, class, wt=N) %>% 
  count(class, stype)

cbp1 <- c("#999999", "#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

#class_type %>%
class_types_sites <-
  SE1 %>%
  filter(event=="asn") %>%
  cs_count(eval_call_srcref, class, stype, wt=N) %>% cs_count(class, stype)

class_types_sites %>%
  ggplot(aes(x=class, y=n, fill=stype)) + 
  geom_bar(stat="identity", position="fill", alpha=0.9) +
  scale_y_continuous(labels=scales::percent) +
#  scale_fill_viridis(discrete = TRUE, option="E") +
  scale_fill_manual(values = cbp1) +
  theme(text = element_text(size=19),
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = .5),
        axis.title.y = element_blank(),
        aspect.ratio = 4/3,
        panel.grid.minor.x = element_blank(),
        legend.position="top") +
  labs(fill="")

ggsave(path(PLOT_DIR, "se-types.pdf"))
```

```{r}
types_sites <- 
  SE1 %>%
  filter(event=="asn") %>%
  cs_count(eval_call_srcref, class, stype, wt=N) %>% cs_count(stype)

closure_sites <- 
  SE1 %>%
  filter(event=="asn") %>%
  cs_count(eval_call_srcref, class, stype, wt=N) %>%
  filter(stype=="Closure") %>% 
  cs_count(class)

overview_table(
  r("Basic type ratio", percent(filter(types_sites, stype == "Basic")$p/100)),
  r("List type ratio", percent(filter(types_sites, stype == "List")$p/100)),
  r("Closure type ratio", percent(filter(types_sites, stype == "Closure")$p/100)),
  r("Closure type", filter(types_sites, stype == "Closure")$n),
  r("Closure type local", filter(closure_sites, class == "Local")$n)
)
```

## Side effects with norm

```{r eval=F}
NE <- read_csv(
  PACKAGE_NORMALIZED_EXPRESSION_FILE,
  col_types = cols(
    minimized = col_character(),
    topcall = col_character(),
    is_model = col_double(),
    has_fundef = col_double(),
    has_calls = col_double(),
    has_assigns = col_double(),
    has_var = col_double(),
    has_bracket = col_double(),
    is_assign = col_double(),
    is_value = col_double(),
    is_ignore = col_double(),
    has_dollar = col_double(),
    has_user_call = col_double(),
    has_block = col_double(),
    has_meta_op = col_double(),
    normalized = col_character(),
    trimmed = col_character(),
    hash = col_character()
  )
) %>%
  select(
    minimized,
    topcall,
    normalized,
    expr_resolved_hash=hash
  )
```

```{r eval=F}
C <- read_fst(path(PACKAGE_DATA_DIR, "calls-sub.fst")) %>% as_tibble
CSE <- left_join(
  C,
  select(SE, cid, event, class, N),
  by="cid"
)
```